---
title: "Myeloid cell reclustering"
author: "Yue (Gary) Zhang"
date: "2023-10-11"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{css, echo=FALSE}
pre {
  max-height: 600px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

# Re-cluster the myeloid cells: cDC, CD16 mono, CD14 mono, pDC


## Preparation {.tabset}

### Load packages

```{r Initiate, echo=FALSE}
## initiate ---------------------------------------------------------
# use the following line to set the library path to /data/zhangy68/R/4.3;
# this path is writtable and used by R in sinteractive mode
# this line is nessacery if running interactive web-based Rstudio
.libPaths('/data/zhangy68/R/4.3')

packages_to_load = c(
    "dplyr",
    "Seurat",
    "patchwork",
    "ggplot2",
    "ggrepel",
    "readxl",
    "future",
    "cowplot",
    "Azimuth",
    "dsb",
    "corrplot",
    "DT",
    "tidyr",
    "ggpubr"
)
# Load packages without printing output
invisible(suppressPackageStartupMessages({
    lapply(packages_to_load, library, character.only = TRUE)
}))

# set variables 
setwd('./')
set.seed('123')
col21 <- rev(c("tomato1","darkblue","turquoise1","lightblue","darkred","mediumblue","purple","bisque",
               "greenyellow","yellow","violetred2","darkgreen","darkgoldenrod1","deeppink3","cadetblue4",
               "orchid2","seagreen3","purple4","dodgerblue2","red","gray27"))

```

### Customized functions
Functions cluster_freq and plot_de_volcano_scatter were created.
```{r function section}
# create a function to generate cluster frequency tables
cluster_freq = function(pbmc,orig.ident,cluster){
  freq <- pbmc@meta.data %>%
    group_by(!!sym(orig.ident), !!sym(cluster)) %>%
    summarise(count=n()) %>%
    mutate(relative_frequency = count/sum(count))
  freq$group = ifelse(freq$orig.ident %in% c('HV-1', 'HV-2', 'HV-3'), 'control', 'trisomy8')
  freq$orig.ident = factor(freq$orig.ident, levels = c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4'))
  return(freq)
}
# custom function to plot DE genes between Trisomy and Normal by clusters
plot_de_volcano_scatter = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Trisomy vs Normal') {
# remove everything after "..." in gene names; bind_rows automatically added these to avoid duplicate rownames.
df$genes = sub(pattern = '\\.\\.\\.*', replacement = '',x = rownames(df))
# calculate log10 q value. cap log10 q value at 5 (which is 0.00001 in q val)
df$p_val_adj_log10 = -log10(df$p_val_adj)
df$p_val_adj_log10_capped = ifelse(df$p_val_adj_log10 < 5, df$p_val_adj_log10, 5)
# generate summary statistics to show on caption
minFDR <- -log10(maxq)
sigs <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) >= minLFC
sigLowFC <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) < minLFC

cnt_plus <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC >= minLFC)
cnt_neg <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC <= -minLFC)
if(showsub) {
  subtitle <- paste0(sum(sigs), " DE genes: ", cnt_plus, " Pos and ", cnt_neg, " Neg",
                     " (FDR<", maxq, ", log2FC>", minLFC, ")")
} else {
  subtitle <- NULL
}

p1 = ggplot(df) +
  geom_point(aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5) + 
  geom_point(data=df[sigs,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigcol) +
  geom_point(data=df[sigLowFC,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigLowFCcol) +
  ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("log10 FDR") +
  facet_wrap(.~Celltype_tested,nrow=round(length(unique(df$Celltype_tested))/2))+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) + 
  geom_vline(xintercept=minLFC, col=sigcol, linetype="dashed", size=0.5) + 
  geom_vline(xintercept=-minLFC, col=sigcol, linetype="dashed", size=0.5) + 
  geom_hline(yintercept=minFDR, col=sigcol, linetype="dashed", size=0.5)

df.s = subset(df, p_val_adj_log10 >minFDR & abs(avg_log2FC) > minLFC)
df.s = df.s[order(df.s$avg_log2FC, decreasing = TRUE), ]
df.s$genes = factor(df.s$genes,levels = unique(df.s$genes))
p2 = ggplot(df.s,mapping = aes(x=genes,y=avg_log2FC,color=p_val_adj_log10_capped)) +
  geom_point() +
  facet_wrap(~ Celltype_tested, scales = 'free') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_flip() +
  ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("genes")
return(list(p1,p2))
}
```

### Load seurat object
load srt.obj.merged.clusterCurated.RDS
```{r load pbmc}
pbmc = readRDS('srt.obj.merged.clusterCuratedRefined.classifiedSet.RDS')
pbmc.copy1 <- pbmc
#pbmc.copy1 -> pbmc

```


### Create a subset of myeloid cells

```{r myeloid subset}
DimPlot(pbmc, reduction = 'wnn.umap', group.by = "curated_clusters3", label = TRUE, label.size = 5, repel = T) + NoLegend()
pbmc.mld = subset(pbmc,curated_clusters3 %in% c('cDC','pDC','CD16 Mono', 'CD14 Mono'))
pbmc.mld@meta.data$orig.ident = factor(pbmc.mld@meta.data$orig.ident, levels = c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4'))
DimPlot(pbmc.mld, reduction = 'wnn.umap', group.by = "curated_clusters3", label = TRUE, label.size = 5, repel = T) + NoLegend()

pbmc.mld.copy <- pbmc.mld
#pbmc.mld.copy -> pbmc.mld

```


## Reclustering the myeloid cells

### Standard preprocessing steps for RNA assay

```{r rna prep}
pbmc.mld.copy -> pbmc.mld2

## RNA assay standard preprocessing steps
pbmc.mld2 <- NormalizeData(pbmc.mld2, normalization.method = "LogNormalize", scale.factor = 10000)
pbmc.mld2 <- FindVariableFeatures(pbmc.mld2, selection.method = "vst", nfeatures = 2000) # find features (genes) expressed most variably between cells
all.genes <- rownames(pbmc.mld2)
pbmc.mld2 <- ScaleData(pbmc.mld2, features = all.genes)
pbmc.mld2 <- RunPCA(pbmc.mld2, features = VariableFeatures(object = pbmc.mld2),reduction.name = 'pca_myeloid',reduction.key = 'pcaMyeloid_')
ElbowPlot(pbmc.mld2,reduction = 'pca_myeloid')
```


### Standard preprocessing steps for ADT assay

```{r adt prep}
# use dsb to correct for background noise and normalize the assay
DefaultAssay(pbmc.mld2) <- 'ADT'
#rownames(pbmc.mld2)

# ADT Isotypes controls (mouse and rat isotypes)
adt.IsotypeControl.list = c("IgG1-1",
                            "IgG2a-1",
                            "IgG2b-1",
                            "IgG1-2",
                            "IgG2a-2",
                            "IgG2b-2",
                            "IgG")
# extract raw ADT count
adt = GetAssayData(pbmc.mld2, slot = 'counts', assay = 'ADT')
# normalize ADTs
dsb.norm = ModelNegativeADTnorm(cell_protein_matrix = adt,
                                denoise.counts = TRUE, 
                                use.isotype.control = TRUE, 
                                isotype.control.name.vec = adt.IsotypeControl.list)

# plot a few proteins to check the results
d = as.data.frame(t(dsb.norm))
par(mfrow = c(4,2))
hist(adt["CD4", ], breaks = 45, col = 'red', main = "CD4", xlab = 'Raw ADT')
hist(d[, "CD4"], breaks = 45, col = '#009ACD80', main = "CD4", xlab = 'ModelNegativeADTnorm')
hist(adt["CD8",], breaks = 45, col = 'red', main = "CD8", xlab = 'Raw ADT')
hist(d[, "CD8"], breaks = 45, col = '#009ACD80', main = "CD8", xlab = 'ModelNegativeADTnorm')
hist(adt["CD3",], breaks = 45, col = 'red', main = "CD3", xlab = 'Raw ADT')
hist(d[, "CD3"], breaks = 45, col = '#009ACD80', main = "CD3", xlab = 'ModelNegativeADTnorm')
hist(adt["CD88",], breaks = 45, col = 'red', main = "CD88", xlab = 'Raw ADT')
hist(d[, "CD88"], breaks = 45, col = '#009ACD80', main = "CD88", xlab = 'ModelNegativeADTnorm')
```

```{r adt prep2}
# add the normalized data to the Seurat object
pbmc.mld2 = SetAssayData(pbmc.mld2, slot = 'data', 
                    assay = 'ADT', 
                    new.data = dsb.norm)
# scale
VariableFeatures(pbmc.mld2) <- rownames(pbmc.mld2[["ADT"]])
pbmc.mld2 <- ScaleData(pbmc.mld2, assay = "ADT")
pbmc.mld2 <- RunPCA(pbmc.mld2, features = rownames(pbmc.mld2), reduction.name = "adt_pca_myeloid", reduction.key = "adtPcaMyeloid_", verbose = FALSE)
DimPlot(pbmc.mld2, reduction = "adt_pca_myeloid")
DimPlot(pbmc.mld2, reduction = "adt_pca")
ElbowPlot(pbmc.mld2,reduction = "adt_pca_myeloid")
```

### WNN clustering

```{r mld2 recluster,fig.width=16, fig.height = 12}
# run WNN 
pbmc.mld2 <- FindMultiModalNeighbors(
  pbmc.mld2, reduction.list = list("pca_myeloid", "adt_pca_myeloid"), 
  dims.list = list(1:20, 1:20), modality.weight.name = c("RNA.weight.myeloid",'ADT.weight.myeloid'), weighted.nn.name = 'weighted.nn.myeloid',
  knn.graph.name = "wknn.myeloid", snn.graph.name = "wsnn.myeloid"
)
pbmc.mld2 <- RunUMAP(pbmc.mld2, nn.name = "weighted.nn.myeloid", 
                reduction.name = "wnn.umap.myeloid", 
                reduction.key = "wnnUMAPmyloid_")
# find clusters with different resolution (0.2 to 1.0)
plot.umap.varyingRes2=NULL
for (i in seq(0.1,1,by=0.1)) {
  pbmc.mld2 <- FindClusters(pbmc.mld2, graph.name = "wsnn.myeloid", 
                       algorithm = 3, resolution = i,
                       verbose = FALSE)
  p1 <- DimPlot(pbmc.mld2, reduction = "wnn.umap.myeloid", label=T) +  NoLegend()
  name.temp = paste0('p',i)
  assign(name.temp,p1)
  print(i)
  print("finished")
}
plot.umap.varyingRes2 = plot_grid(p0.1,p0.2,p0.3,p0.4,p0.5,p0.6,p0.7,p0.8,p0.9,p1, labels = c(paste0('res=', seq(0.1,1,by=0.1))), ncol=5)
plot.umap.varyingRes2
```

### Review clusters

The Dimplot shows: 
1. The controls (first line in the split dimplot) are relatively more similar to each other in comparison with the patient group
2. On the group-level comparison, patients have more cells in cluster 1, especially in TRIAD-3, TRIAD-7 and TRIAD-4.
3. TRIAD-1 has an distinctly dense cluster 2, TRIAD-7 has substantially more cells in clusetrs 1 and 4, TRIAD-2 has a distinctly dense cluster 3

```{r curate clusters,fig.width=16, fig.height = 12}
Idents(pbmc.mld2) <- "wsnn.myeloid_res.0.2" 
DimPlot(pbmc.mld2, reduction = "wnn.umap.myeloid", label=T)
DimPlot(pbmc.mld2, reduction = "wnn.umap.myeloid", split.by = 'wsnn.myeloid_res.0.2', ncol = 3, label=T)
DimPlot(pbmc.mld2, reduction = "wnn.umap.myeloid", split.by = 'orig.ident', ncol = 3, label=T)
```

### Cluster frequency calculation

```{r cluster freq,fig.width=5, fig.height = 5}

pbmc.mld2@meta.data$myeloid_Clusters = pbmc.mld2@meta.data$wsnn.myeloid_res.0.2
pbmc.mld2.cluFreq = cluster_freq(pbmc.mld2,'orig.ident','myeloid_Clusters')
ggplot(data = pbmc.mld2.cluFreq, mapping = aes(x = orig.ident, y = `relative_frequency`,fill = myeloid_Clusters)) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Myeloid cells cluster frequency') +
  guides(fill=guide_legend(nrow=2))
```


## Save and show session info {.tabset}

### Save the seurat object as .rds files
```{r save}
saveRDS(pbmc.mld2,'pbmc.mld2.RDS')
```

### Session info
```{r sessioninfo}
sessionInfo()
```