---
title: "Myeloid cell cluster analysis"
date: "`r Sys.Date()`"
author: Yue (Gary) Zhang, PhD
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Preparation

## Load packages and set variables

```{r Initiate, results='hide'}
## initiate ---------------------------------------------------------
# use the following line to set the library path to /data/zhangy68/R/4.3;
# this path is writtable and used by R in sinteractive mode
# this line is nessacery if running interactive web-based Rstudio
.libPaths('/data/zhangy68/R/4.3')

packages_to_load = c(
    "dplyr",
    "Seurat",
    "patchwork",
    "ggplot2",
    "ggrepel",
    "readxl",
    "future",
    "cowplot",
    "Azimuth",
    "dsb",
    "corrplot",
    "DT",
    "tidyr",
    "ggpubr",
    "scRNAtoolVis"
)
# Load packages without printing output
invisible(suppressPackageStartupMessages({
    lapply(packages_to_load, library, character.only = TRUE)
}))

# set variables 
setwd('./')
set.seed('123')
col21 <- rev(c("tomato1","darkblue","turquoise1","lightblue","darkred","mediumblue","purple","bisque",
               "greenyellow","yellow","violetred2","darkgreen","darkgoldenrod1","deeppink3","cadetblue4",
               "orchid2","seagreen3","purple4","dodgerblue2","red","gray27"))

```

## Customized functions

```{r function section}
# create a function to generate cluster frequency tables
cluster_freq = function(pbmc,orig.ident,cluster){
  freq <- pbmc@meta.data %>%
    group_by(!!sym(orig.ident), !!sym(cluster)) %>%
    summarise(count=n()) %>%
    mutate(relative_frequency = count/sum(count))
  freq$group = ifelse(freq$orig.ident %in% c('HV-1', 'HV-2', 'HV-3'), 'control', 'trisomy8')
  freq$orig.ident = factor(freq$orig.ident, levels = c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4'))
  return(freq)
}
# custom function to plot DE genes between Trisomy and Normal by clusters
plot_de_volcano_scatter = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Trisomy vs Normal') {
# remove everything after "..." in gene names; bind_rows automatically added these to avoid duplicate rownames.
df$genes = sub(pattern = '\\.\\.\\.*', replacement = '',x = rownames(df))
# calculate log10 q value. cap log10 q value at 5 (which is 0.00001 in q val)
df$p_val_adj_log10 = -log10(df$p_val_adj)
df$p_val_adj_log10_capped = ifelse(df$p_val_adj_log10 < 5, df$p_val_adj_log10, 5)
# generate summary statistics to show on caption
minFDR <- -log10(maxq)
sigs <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) >= minLFC
sigLowFC <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) < minLFC

cnt_plus <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC >= minLFC)
cnt_neg <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC <= -minLFC)
if(showsub) {
  subtitle <- paste0(sum(sigs), " DE genes: ", cnt_plus, " Pos and ", cnt_neg, " Neg",
                     " (FDR<", maxq, ", log2FC>", minLFC, ")")
} else {
  subtitle <- NULL
}

p1 = ggplot(df) +
  geom_point(aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5) + 
  geom_point(data=df[sigs,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigcol) +
  geom_point(data=df[sigLowFC,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigLowFCcol) +
  ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("log10 FDR") +
  facet_wrap(.~Celltype_tested,nrow=round(length(unique(df$Celltype_tested))/2))+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) + 
  geom_vline(xintercept=minLFC, col=sigcol, linetype="dashed", size=0.5) + 
  geom_vline(xintercept=-minLFC, col=sigcol, linetype="dashed", size=0.5) + 
  geom_hline(yintercept=minFDR, col=sigcol, linetype="dashed", size=0.5)

df.s = subset(df, p_val_adj_log10 >minFDR & abs(avg_log2FC) > minLFC)
df.s = df.s[order(df.s$avg_log2FC, decreasing = TRUE), ]
df.s$genes = factor(df.s$genes,levels = unique(df.s$genes))
p2 = ggplot(df.s,mapping = aes(x=genes,y=avg_log2FC,color=p_val_adj_log10_capped)) +
  geom_point() +
  facet_wrap(~ Celltype_tested, scales = 'free') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_flip() +
  ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("genes")
return(list(p1,p2))
}

create_dt <- function(df,caption){
  datatable(df,
                caption = caption,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))))
}
  
```

## Load seurat object

Load the seurat object of re-clustered myeloid cells generated in Session 14 (pbmc.mld2.RDS).

```{r load pbmc}
pbmc.mld2 = readRDS('pbmc.mld2.RDS')
pbmc.mld2.copy1 <- pbmc.mld2
#pbmc.mld2.copy1 -> pbmc.mld2
Idents(pbmc.mld2) = pbmc.mld2@meta.data$myeloid_Clusters
```

# UMAP Dimplots

## All samples combined

```{r dim1, fig.height= 6, fig.width= 6}
DimPlot(pbmc.mld2, reduction = "wnn.umap.myeloid", label=T)
```

## Myeloid cell cluster frequency

```{r mc,fig.width= 8, fig.height= 8}
pbmc.mld2.cluFreq = cluster_freq(pbmc.mld2,'orig.ident','myeloid_Clusters')
ggplot(data = pbmc.mld2.cluFreq, mapping = aes(x = orig.ident, y = `relative_frequency`,fill = myeloid_Clusters)) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Myeloid cells cluster frequency') +
  guides(fill=guide_legend(nrow=2))

```

# Differential Expression Analysis Across All Myeloid Cell Subclusters

Use findAllMarkers to recursively compare gene/ADT expression between each cluster and all the rest clusters combined. Only genes/ADTs expressed in at least 50% of cells in either populations were tested, and only genes/ADTs with q<0.1 and the absolute values of lfc > 0.25 were shown. 

## RNA markers

```{r findallmarkers rna, results='hide'}
DefaultAssay(pbmc.mld2) = 'RNA'
cluster.markers.RNA = FindAllMarkers(pbmc.mld2,assay = 'RNA',min.pct = 0.5, only.pos = F, logfc.threshold = 0.25)
```

```{r show signature rna, fig.height= 12, fig.width= 16}
# parse the result table: 
# 1. select rows with q < 0.1; 
# 2. rank rows within each cluster by descending lfc
# 3. reorder the columns 

cluster.markers.RNA.sig = subset(cluster.markers.RNA,p_val_adj < 0.1)
cluster.markers.RNA.sig <- cluster.markers.RNA.sig %>%
  arrange(cluster, desc(avg_log2FC)) %>%  # Arrange by cluster and descending lfc
  group_by(cluster)              # Group by cluster
cluster.markers.RNA.sig = cluster.markers.RNA.sig[,c(6,7,3,4,2,1,5)]
# show the full parsed table
create_dt(cluster.markers.RNA.sig,caption = 'Signature genes of each cluster (q < 0.1, ranked by lfc) of each cluster')
```

```{r heatmap RNA, fig.height= 12, fig.width= 16, results='hide'}
# select the top 5 genes for each cluster and plot them
cluster.markers.RNA.sig.top5 <-cluster.markers.RNA.sig %>%
  arrange(cluster, desc(avg_log2FC)) %>%  # Arrange by cluster and descending lfc
  group_by(cluster) %>%                # Group by cluster
  slice_head(n = 5)  # only show the top 5 genes for each cluster

AverageHeatmap(object = pbmc.mld2,
               group.by = 'myeloid_Clusters',
               assays = 'RNA',
               markerGene = cluster.markers.RNA.sig.top5$gene,
               row_title = 'Cluster Top 5 Marker Genes')
```
