---
title: "Trisomy/Disomy cell frequency analysis"
date: "`r Sys.Date()`"
author: Yue (Gary) Zhang, PhD
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Preparation

## Load packages and set variables

```{r Initiate, results='hide'}
## initiate ---------------------------------------------------------
# use the following line to set the library path to /data/zhangy68/R/4.3;
# this path is writtable and used by R in sinteractive mode
# this line is nessacery if running interactive web-based Rstudio
.libPaths('/data/zhangy68/R/4.3')

packages_to_load = c(
    "dplyr",
    "Seurat",
    "patchwork",
    "ggplot2",
    "ggrepel",
    "readxl",
    "future",
    "cowplot",
    "Azimuth",
    "dsb",
    "corrplot",
    "DT",
    "tidyr",
    "ggpubr",
    "scRNAtoolVis"
)
# Load packages without printing output
invisible(suppressPackageStartupMessages({
    lapply(packages_to_load, library, character.only = TRUE)
}))

# set variables 
setwd('./')
set.seed('123')
col21 <- rev(c("tomato1","darkblue","turquoise1","lightblue","darkred","mediumblue","purple","bisque",
               "greenyellow","yellow","violetred2","darkgreen","darkgoldenrod1","deeppink3","cadetblue4",
               "orchid2","seagreen3","purple4","dodgerblue2","red","gray27"))

```

## Customized functions

```{r function section}
# create a function to generate cluster frequency tables
cluster_freq = function(pbmc,orig.ident,cluster){
  freq <- pbmc@meta.data %>%
    group_by(!!sym(orig.ident), !!sym(cluster)) %>%
    summarise(count=n()) %>%
    mutate(relative_frequency = count/sum(count))
  freq$group = ifelse(freq$orig.ident %in% c('HV-1', 'HV-2', 'HV-3'), 'control', 'trisomy8')
  freq$orig.ident = factor(freq$orig.ident, levels = c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4'))
  return(freq)
}
# custom function to plot DE genes between Trisomy and Normal by clusters
plot_de_volcano_scatter = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Trisomy vs Normal') {
# remove everything after "..." in gene names; bind_rows automatically added these to avoid duplicate rownames.
df$genes = sub(pattern = '\\.\\.\\.*', replacement = '',x = rownames(df))
# calculate log10 q value. cap log10 q value at 5 (which is 0.00001 in q val)
df$p_val_adj_log10 = -log10(df$p_val_adj)
df$p_val_adj_log10_capped = ifelse(df$p_val_adj_log10 < 5, df$p_val_adj_log10, 5)
# generate summary statistics to show on caption
minFDR <- -log10(maxq)
sigs <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) >= minLFC
sigLowFC <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) < minLFC

cnt_plus <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC >= minLFC)
cnt_neg <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC <= -minLFC)
if(showsub) {
  subtitle <- paste0(sum(sigs), " DE genes: ", cnt_plus, " Pos and ", cnt_neg, " Neg",
                     " (FDR<", maxq, ", log2FC>", minLFC, ")")
} else {
  subtitle <- NULL
}

p1 = ggplot(df) +
  geom_point(aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5) + 
  geom_point(data=df[sigs,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigcol) +
  geom_point(data=df[sigLowFC,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigLowFCcol) +
  ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("log10 FDR") +
  facet_wrap(.~Celltype_tested,nrow=round(length(unique(df$Celltype_tested))/2))+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) + 
  geom_vline(xintercept=minLFC, col=sigcol, linetype="dashed", size=0.5) + 
  geom_vline(xintercept=-minLFC, col=sigcol, linetype="dashed", size=0.5) + 
  geom_hline(yintercept=minFDR, col=sigcol, linetype="dashed", size=0.5)

df.s = subset(df, p_val_adj_log10 >minFDR & abs(avg_log2FC) > minLFC)
df.s = df.s[order(df.s$avg_log2FC, decreasing = TRUE), ]
df.s$genes = factor(df.s$genes,levels = unique(df.s$genes))
p2 = ggplot(df.s,mapping = aes(x=genes,y=avg_log2FC,color=p_val_adj_log10_capped)) +
  geom_point() +
  facet_wrap(~ Celltype_tested, scales = 'free') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_flip() +
  ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("genes")
return(list(p1,p2))
}

create_dt <- function(df,caption){
  datatable(df,
                caption = caption,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))))
}
  
```

## Load seurat object

load srt.obj.merged.clusterCurated.RDS

```{r load pbmc}
pbmc = readRDS('srt.obj.merged.clusterCuratedRefined.classifiedSet.RDS')
pbmc.copy1 <- pbmc
#pbmc.copy1 -> pbmc
```

## Attach trisomy/normal celltyping information

```{r attach trisomy-normal info, results='hide'}
tsct = read.csv('Trisomy-Annotation-Shouguo-May-2023.merged.csv')
# parse celll barcode to match metadata
tsct$cell = sub('sample','',tsct$cell)
# most of cell barcodes have "normal" as prefix; they should be redundant info; remove them
tsct = subset(tsct,!grepl(pattern = 'normal',x = tsct$cell))
# Check how many cell barcodes share between seurat and tsct 
length(intersect(rownames(pbmc@meta.data),tsct$cell)) # 55693
nrow(pbmc@meta.data) # 83107
# merge tsct with metadata
pbmc@meta.data = merge(pbmc@meta.data,select(tsct,c(cell,RatioChr8,trisomy)),
                 by.x = 0, by.y = 'cell', all.x = T)
# sanity check
nrow(pbmc@meta.data[complete.cases(pbmc@meta.data$trisomy),]) # 55693; as same as the number of shared cells between sheets above
rownames(pbmc@meta.data) = pbmc@meta.data$Row.names
#pbmc@meta.data = select(pbmc@meta.data, -c(RatioChr8.x,trisomy.x,RatioChr8.y,trisomy.y))

```

# Trisomy/Normal Celltyping Analysis (All Cells)

## Trisomy/Normal for each patient (without cell cluster annotation)

### Summary statistics chart

```{r ct, results='hide'}
tsct.summary = cluster_freq(pbmc,'orig.ident','trisomy')
tsct.summary$trisomy[is.na(tsct.summary$trisomy)] = 'NA in Trisomy annotation'
sum(tsct.summary$count) # 83107
sum(tsct.summary$count[tsct.summary$trisomy != 'NA in Trisomy annotation']) # 55693; match to the value above
# write.csv(tsct.summary, 'tsct.summary.csv')
```

```{r show table}
create_dt(tsct.summary,caption = 'Trisomy vs Normal per Patient')
```

### Stacked barplots: the number of cells annotated

```{r cell typing count}
ggplot(data = tsct.summary, mapping = aes(x = orig.ident, y = `count`,fill = trisomy)) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Cell Count') +
  guides(fill=guide_legend(nrow=2))
  
```

### Stacked barplots: the frequency of cells annotated

```{r plot frequency}
ggplot(data = tsct.summary, mapping = aes(x = orig.ident, y = `relative_frequency`,fill = trisomy)) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Celltyping Frequency') +
  guides(fill=guide_legend(nrow=2))
```


## Trisomy/Normal in each cell cluster

### Summary statistics chart

```{r analyze cell typing by clusters, results='hide'}
pbmc@meta.data %>%
    group_by(orig.ident, curated_clusters4, trisomy) %>%
    summarise(count=n()) %>%
    mutate(relative_frequenc = count/sum(count)) -> tsct.summary2
tsct.summary2$trisomy[is.na(tsct.summary2$trisomy)] = 'NA in Trisomy annotation'
# write.csv(tsct.summary2, 'tsct.summary2.csv')
sum(tsct.summary2$count) # 83107
sum(tsct.summary2$count[tsct.summary2$trisomy != 'NA in Trisomy annotation']) # 55693; match to the value above
sum(tsct.summary2$count[tsct.summary2$orig.ident == 'TRIAD-1']) == nrow(pbmc@meta.data[pbmc@meta.data$orig.ident=='TRIAD-1',]) # matched
tsct.summary3 <- pivot_wider(select(tsct.summary2, - relative_frequenc), names_from = trisomy, values_from = count, values_fill = 0) %>%
  subset(!(orig.ident %in% c('HV-1', 'HV-2', 'HV-3')))
```

```{r freq cl}
create_dt(tsct.summary3,caption = 'Trisomy vs Normal in each cell cluster per Patient')
```

### Stacked barplots: the number of cells annotated in each cluster

Trisomy vs Normal; gray and unannotated cells included

```{r cn by cluster, fig.width=15}
ggplot(data = tsct.summary2[!(tsct.summary2$orig.ident %in% c('HV-1','HV-2','HV-3')),], mapping = aes(x = curated_clusters4, y = `count`,fill = trisomy)) +
  facet_grid(.~orig.ident) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Cell Count') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal', subtitle = 'Gray and unannotated cells included')
```

### Stacked barplots: the frequency of cells annotated in each cluster

Trisomy vs Normal; gray and unannotated cells included

```{r freq by cluster, fig.width=15}
ggplot(data = tsct.summary2[!(tsct.summary2$orig.ident %in% c('HV-1','HV-2','HV-3')),], mapping = aes(x = curated_clusters4, y = `relative_frequenc`,fill = trisomy)) +
  facet_grid(.~orig.ident) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Frequency') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal', subtitle = 'Gray and unannotated cells included')
```

Trisomy vs Normal; gray and unannotated cells excluded
```{r freq by cluster2, fig.width=15}
ggplot(data = tsct.summary2 %>%
         subset(!(orig.ident %in% c('HV-1','HV-2','HV-3'))) %>%
         subset(trisomy %in% c('NORMAL','TRISOMY')),, mapping = aes(x = curated_clusters4, y = `count`,fill = trisomy)) +
  facet_grid(.~orig.ident) +
  geom_bar(stat="identity",color = 'black',position = 'fill') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Frequency') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal', subtitle = 'Gray and unannotated cells excluded')
```

Trisomy vs Normal; gray and unannotated cells excluded; grouped by cell clusters

```{r freq by cluster3, fig.width=15}
ggplot(data = tsct.summary2 %>%
         subset(!(orig.ident %in% c('HV-1','HV-2','HV-3'))) %>%
         subset(trisomy %in% c('NORMAL','TRISOMY')),, mapping = aes(x = orig.ident, y = `count`,fill = trisomy)) +
  facet_grid(.~curated_clusters4) +
  geom_bar(stat="identity",color = 'black',position = 'fill') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 9, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Frequency') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal', subtitle = 'Gray and unannotated cells excluded; grouped by cell clusters')

```

### Barplot: The average of the trisomy cell percentage for each cell cluster

Here we calculate the average and standard deviation of the percentage of trisomy cells (trisomy/trisomy+normal) across all patients for each cell cluster.

```{r freq avg}
tsct.summary4 = tsct.summary2 %>%
  subset(trisomy == 'TRISOMY', select = c(curated_clusters4,relative_frequenc)) %>%
  group_by(curated_clusters4) %>%
  summarize(
    average_frequency = mean(relative_frequenc),
    std_frequency = sd(relative_frequenc)
  )

# No Trisomy cells in Plasmablast B cells at all, so let's manually add it as 0 in the table for plotting
tsct.summary4 = rbind(tsct.summary4, c('Plasmablast B', 0,0))
tsct.summary4$average_frequency = as.numeric(tsct.summary4$average_frequency)
tsct.summary4$std_frequency = as.numeric(tsct.summary4$std_frequency)

tsct.summary4$curated_clusters4 = factor(tsct.summary4$curated_clusters4,c( "CD14 Mono", 
                                                                            "CD16 Mono",  
                                                                            "cDC", 
                                                                            "pDC", 
                                                                            "B Naive", 
                                                                            "B memory",
                                                                            'Plasmablast B',
                                                                           "CD4 T Naive", "CD4 T Memory",
                                                                           "CD8 T Naive", "CD8 T Memory",
                                                                            "NK", 
                                                                           "MAIT","gdT"))
ggplot(tsct.summary4, aes(x = curated_clusters4, y = average_frequency)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_errorbar(aes(ymin = average_frequency - std_frequency,
                    ymax = average_frequency + std_frequency),
                width = 0.2, position = position_dodge(0.9)) +
  labs(title = "The fraction of trisomy cells in cell clusters across all patients",
       x = "Cell clusters",
       y = "Fraction of Trisomy cells") +
theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=12,face="bold"))

```

# Trisomy/Normal Celltyping Analysis (Myeloid Cells)

## Preprocessing

Load the seurat object of re-clustered myeloid cells generated in Session 14 (pbmc.mld2.RDS).
Merge the myeloid cell subcluster index to metadata.

```{r load pbmcmld2, results='hide'}
pbmc.mld2 = readRDS('pbmc.mld2.RDS')
pbmc.mld2.copy1 <- pbmc.mld2
#pbmc.mld2.copy1 -> pbmc.mld2
pbmc.mld2@meta.data %>%
  as.data.frame() %>%
  select(myeloid_Clusters) %>%
  merge(pbmc@meta.data,.,
                 by = 0, all.x = T) -> temp
nrow(temp) # 83107; as same as above
nrow(subset(temp,!is.na(myeloid_Clusters))) == nrow(pbmc.mld2@meta.data) # TRUE; it is 20910 cell; merging is successful
all(temp[,1] == temp[,2])
pbmc@meta.data <- temp[,-2] # remove the redundant column
rownames(pbmc@meta.data) = pbmc@meta.data$Row.name
```


## Summary statistics chart

```{r analyze cell typing by clusters mld, results='hide'}
pbmc@meta.data %>%
    group_by(orig.ident, myeloid_Clusters, trisomy) %>%
    summarise(count=n()) %>%
    mutate(relative_frequenc = count/sum(count)) -> tsct.mld.summary2
tsct.mld.summary2$trisomy[is.na(tsct.mld.summary2$trisomy)] = 'NA in Trisomy annotation'
# write.csv(tsct.mld.summary2, 'tsct.mld.summary2.csv')
sum(tsct.mld.summary2$count) # 83107
sum(tsct.mld.summary2$count[tsct.mld.summary2$trisomy != 'NA in Trisomy annotation']) # 55693; match to the value above
sum(tsct.mld.summary2$count[tsct.mld.summary2$orig.ident == 'TRIAD-1']) == nrow(pbmc@meta.data[pbmc@meta.data$orig.ident=='TRIAD-1',]) # matched
sum(tsct.mld.summary2$count[tsct.mld.summary2$orig.ident == 'TRIAD-1' & !(is.na(tsct.mld.summary2$myeloid_Clusters))]) == nrow(pbmc.mld2@meta.data[pbmc.mld2@meta.data$orig.ident=='TRIAD-1',]) # matched
tsct.mld.summary3 <- pivot_wider(select(tsct.mld.summary2, - relative_frequenc), names_from = trisomy, values_from = count, values_fill = 0) %>%
  subset(!(orig.ident %in% c('HV-1', 'HV-2', 'HV-3')))
```

```{r freq cl mld}
create_dt(tsct.mld.summary3,caption = 'Trisomy vs Normal in each myeloid cell subcluster per Patient')
```

## Stacked barplots: the number of cells annotated in each cluster

Trisomy vs Normal in Myeloid Cell Subclusters; gray and unannotated cells included
```{r frequency by cluster mld, fig.width=15}
# filter out information with myeloid cell sub-clusters
tsct.mld.summary2 = tsct.mld.summary2[complete.cases(tsct.mld.summary2$myeloid_Clusters),] 
ggplot(data = tsct.mld.summary2[!(tsct.mld.summary2$orig.ident %in% c('HV-1','HV-2','HV-3')),], mapping = aes(x = myeloid_Clusters, y = `count`,fill = trisomy)) +
  facet_grid(.~orig.ident) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
  xlab("Subject") + ylab('Cell Count') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal in Myeloid Cell Subclusters', subtitle = 'Gray and unannotated cells included')
```

Trisomy vs Normal in Myeloid Cell Subclusters; gray and unannotated cells included; grouped by cell clusters

```{r frequency by cluster mld 2, fig.width=15}
ggplot(data = tsct.mld.summary2[!(tsct.mld.summary2$orig.ident %in% c('HV-1','HV-2','HV-3')),], mapping = aes(x = orig.ident, y = `count`,fill = trisomy)) +
  facet_grid(.~myeloid_Clusters) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
  xlab("Subject") + ylab('Cell Count') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal in Myeloid Cell Subclusters', subtitle = 'Gray and unannotated cells included; grouped by cell clusters')
```

## Stacked barplots: the frequency of cells annotated in each cluster

Trisomy vs Normal in Myeloid Cell Subclusters; gray and unannotated cells included

```{r frequency by cluster mld 3, fig.width=15}
ggplot(data = tsct.mld.summary2[!(tsct.mld.summary2$orig.ident %in% c('HV-1','HV-2','HV-3')),], mapping = aes(x = myeloid_Clusters, y = `relative_frequenc`,fill = trisomy)) +
  facet_grid(.~orig.ident) +
  geom_bar(stat="identity",color = 'black') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
  xlab("Subject") + ylab('Frequency') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal in Myeloid Cell Subclusters', subtitle = 'Gray and unannotated cells included')
```

Trisomy vs Normal in Myeloid Cell Subclusters; gray and unannotated cells excluded

```{r frequency by cluster mld2 4, fig.width=15}
ggplot(data = tsct.mld.summary2 %>%
         subset(!(orig.ident %in% c('HV-1','HV-2','HV-3'))) %>%
         subset(trisomy %in% c('NORMAL','TRISOMY')), mapping = aes(x = myeloid_Clusters, y = count,fill = trisomy)) +
  facet_grid(.~orig.ident) +
  geom_bar(stat="identity",color = 'black',position = 'fill') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
  xlab("Subject") + ylab('Frequency') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal in Myeloid Cell Subclusters', subtitle = 'Gray and unannotated cells excluded')
```

Trisomy vs Normal in Myeloid Cell Subclusters; gray and unannotated cells excluded; grouped by cell clusters

```{r frequency by cluster mld 5, fig.width=15}
ggplot(data = tsct.mld.summary2 %>%
         subset(!(orig.ident %in% c('HV-1','HV-2','HV-3'))) %>%
         subset(trisomy %in% c('NORMAL','TRISOMY')),, mapping = aes(x = orig.ident, y = `count`,fill = trisomy)) +
  facet_grid(.~myeloid_Clusters) +
  geom_bar(stat="identity",color = 'black',position = 'fill') +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 9, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
  xlab("Subject") + ylab('Frequency') +
  guides(fill=guide_legend(nrow=2)) + ggtitle('Trisomy vs Normal in Myeloid Cell Subclusters', subtitle = 'Gray and unannotated cells excluded; grouped by cell clusters')

```

## Barplot: The average of the trisomy cell percentage in myeloid cell subclusters

Here we calculate the average and standard deviation of the percentage of trisomy cells (trisomy/trisomy+normal) across all patients for each myeloid cell subcluster.

```{r freq mld avg}
tsct.mld.summary4 = tsct.mld.summary2 %>%
  subset(trisomy == 'TRISOMY', select = c(myeloid_Clusters,relative_frequenc)) %>%
  group_by(myeloid_Clusters) %>%
  summarize(
    average_frequency = mean(relative_frequenc),
    std_frequency = sd(relative_frequenc)
  )

ggplot(tsct.mld.summary4, aes(x = myeloid_Clusters, y = average_frequency)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_errorbar(aes(ymin = average_frequency - std_frequency,
                    ymax = average_frequency + std_frequency),
                width = 0.2, position = position_dodge(0.9)) +
  labs(title = "The fraction of trisomy cells in myeloid cell subclusters across all patients",
       x = "Myeloid cell subclusters",
       y = "Trisomy/Normal cell counts") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=12,face="bold"))

```

## Save the seurat object as .rds files
```{r save}
saveRDS(pbmc,'srt.obj.merged.clusterCuratedRefined.classifiedSet.myeloidInfoAdded.RDS')
```
