---
title: "Functional score, Cell cluster frequency plots"
date: "`r Sys.Date()`"
author: Yue (Gary) Zhang, PhD
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = T, message = FALSE) 
```

# Preparation

## Load packages and set variables

```{r Initiate, results='hide'}
# use the following line to set the library path to /data/zhangy68/R/4.3;
# this path is writtable and used by R in sinteractive mode
# this line is nessacery if running interactive web-based Rstudio
.libPaths('/data/zhangy68/R/4.3')

packages_to_load = c(
  "dplyr",
  "Seurat",
  "patchwork",
  "ggplot2",
  "ggrepel",
  "readxl",
  "future",
  "cowplot",
  "Azimuth",
  "dsb",
  "corrplot",
  "DT",
  "tidyr",
  "ggpubr",
  "scRNAtoolVis",
  "purrr",
  "limma",
  "wesanderson",
  "reshape",
  "edgeR",
  'ggfun',
  'ggrepel',
  'fgsea',
  'circlize',
  'ComplexHeatmap',
  'vegan',
  "immunarch"
)
# Load packages without printing output
invisible(suppressPackageStartupMessages({
  lapply(packages_to_load, library, character.only = TRUE)
}))

# Immcantation required packages
suppressPackageStartupMessages(library(airr))
suppressPackageStartupMessages(library(alakazam))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(dowser))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(scoper))
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(shazam))

# set variables 
setwd('./')
set.seed('123')
col21 <- rev(c("tomato1","darkblue","turquoise1","lightblue","darkred","mediumblue","purple","bisque",
               "greenyellow","yellow","violetred2","darkgreen","darkgoldenrod1","deeppink3","cadetblue4",
               "orchid2","seagreen3","purple4","dodgerblue2","red","gray27"))

```

## Customized functions

```{r function section}
# create a function to generate cluster frequency tables
cluster_freq = function(pbmc,orig.ident,cluster,sex=F){
  freq <- pbmc@meta.data %>%
    group_by(!!sym(orig.ident), !!sym(cluster)) %>%
    summarise(count=n()) %>%
    mutate(relative_frequency = count/sum(count))
  freq$group = ifelse(freq$orig.ident %in% c('HV-1', 'HV-2', 'HV-3'), 'control', 'trisomy8')
  if(sex == T){
    freq$sex = cov$sex[match(freq$orig.ident,cov$sample)]
    freq$group_sex = paste0(freq$group,'_',freq$sex)
    freq$group_sex = factor(freq$group_sex, levels = c('control_F','trisomy8_F','control_M','trisomy8_M'))
  }
  freq$orig.ident = factor(freq$orig.ident, levels = c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4'))
  return(freq)
}
# custom function to plot DE genes between Trisomy and Normal by clusters
plot_de_volcano_scatter = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Trisomy vs Normal') {
  # remove everything after "..." in gene names; bind_rows automatically added these to avoid duplicate rownames.
  df$genes = sub(pattern = '\\.\\.\\.*', replacement = '',x = rownames(df))
  # calculate log10 q value. cap log10 q value at 5 (which is 0.00001 in q val)
  df$p_val_adj_log10 = -log10(df$p_val_adj)
  df$p_val_adj_log10_capped = ifelse(df$p_val_adj_log10 < 5, df$p_val_adj_log10, 5)
  # generate summary statistics to show on caption
  minFDR <- -log10(maxq)
  sigs <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) >= minLFC
  sigLowFC <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) < minLFC
  
  cnt_plus <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC >= minLFC)
  cnt_neg <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC <= -minLFC)
  if(showsub) {
    subtitle <- paste0(sum(sigs), " DE genes: ", cnt_plus, " Pos and ", cnt_neg, " Neg",
                       " (FDR<", maxq, ", log2FC>", minLFC, ")")
  } else {
    subtitle <- NULL
  }
  
  p1 = ggplot(df) +
    geom_point(aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5) + 
    geom_point(data=df[sigs,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigcol) +
    geom_point(data=df[sigLowFC,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigLowFCcol) +
    ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("log10 FDR") +
    facet_wrap(.~Celltype_tested,nrow=round(length(unique(df$Celltype_tested))/2))+
    theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) + 
    geom_vline(xintercept=minLFC, col=sigcol, linetype="dashed", size=0.5) + 
    geom_vline(xintercept=-minLFC, col=sigcol, linetype="dashed", size=0.5) + 
    geom_hline(yintercept=minFDR, col=sigcol, linetype="dashed", size=0.5)
  
  df.s = subset(df, p_val_adj_log10 >minFDR & abs(avg_log2FC) > minLFC)
  df.s = df.s[order(df.s$avg_log2FC, decreasing = TRUE), ]
  df.s$genes = factor(df.s$genes,levels = unique(df.s$genes))
  p2 = ggplot(df.s,mapping = aes(x=genes,y=avg_log2FC,color=p_val_adj_log10_capped)) +
    geom_point() +
    facet_wrap(~ Celltype_tested, scales = 'free') +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    coord_flip() +
    ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("genes")
  return(list(p1,p2))
}

create_dt <- function(df,caption){
  datatable(df,
            caption = caption,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All")),
                           initComplete = JS(
    "function(settings, json) {",
    "$('caption').css({'font-size': '20px', 'font-family': 'Arial', 'font-weight': 'bold'});",
    "}")))
}

# function to set a column as rownames and remove that column from the df; default is setting the first column as the rowname
setrowname =  function(df,col = 1){
  if (is.numeric(col)) {
    # print('it is a number')
    rownames(df) = df[,col]
    df = df[,-col]
  } else {
    # print('it is a string')
    rownames(df) = df[,which(names(df) %in% c(col))]
    df = df[,-which(names(df) %in% c(col))]
  }
  return(df)
} 

# function to get colors from a palatte
getcol = function(group,pal = "Darjeeling1",shift_col_palatte=0){
  groupcol = as.numeric(factor(group))
  groupcol = wes_palette(pal)[groupcol+shift_col_palatte]
  return(groupcol)
}


# Volcano plot
plot_de_volcano_scatter3 = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Volcano plot',
                                    axislim,leftarrowlim,rightarrowlim) {
  # calculate log10 q value. 
  df$qval_log10 = -log10(df$`adj.P.Val`)
  # generate summary statistics to show on caption
  minFDR <- -log10(maxq)
  sigs <- df$qval_log10 >= minFDR & abs(df$logFC) >= minLFC
  sigLowFC <- df$qval_log10 >= minFDR & abs(df$logFC) < minLFC
  
  cnt_plus <- sum(df$qval_log10 >= minFDR & df$logFC >= minLFC)
  cnt_neg <- sum(df$qval_log10 >= minFDR & df$logFC <= -minLFC)
  if(showsub) {
    subtitle <- paste0(sum(sigs), " DEGs: ", cnt_plus, " + and ", cnt_neg, " -",
                       " (q<", maxq, ", log2FC>", minLFC, ")")
  } else {
    subtitle <- NULL
  }
  
  p1 = ggplot(df) +
    geom_point(aes(x=logFC, y=qval_log10, color = logFC, size = qval_log10)) + 
    geom_point(data =  df %>%
                 subset(qval_log10 >= minFDR) %>%
                 dplyr::arrange(desc(logFC)) %>%
                 dplyr::slice(1:5),
               aes(x = logFC, y = qval_log10,
                   # fill = log2FoldChange,
                   size = qval_log10),
               shape = 21, show.legend = F, color = "#000000") +
    geom_text_repel(data =  df %>%
                      subset(qval_log10 >= minFDR) %>%
                      dplyr::arrange(desc(logFC)) %>%
                      dplyr::slice(1:5),
                    aes(x = logFC, y = qval_log10,label = geneID),
                    box.padding = 0.5,
                    nudge_x = 0.5,
                    nudge_y = 0.2,
                    segment.curvature = -0.1,
                    segment.ncp = 3,
                    # segment.angle = 10,
                    direction = "y", 
                    hjust = "left"
    ) + 
    geom_point(data =  df %>%
                 subset(qval_log10 >= minFDR) %>%
                 dplyr::arrange(logFC) %>%
                 dplyr::slice(1:5),
               aes(x = logFC, y = qval_log10,
                   # fill = log2FoldChange,
                   size = qval_log10),
               shape = 21, show.legend = F, color = "#000000") +
    geom_text_repel(data =  df %>%
                      subset(qval_log10 >= minFDR) %>%
                      dplyr::arrange(logFC) %>%
                      dplyr::slice(1:5),
                    aes(x = logFC, y = qval_log10,label = geneID),
                    box.padding = 0.5,
                    nudge_x = 0.5,
                    nudge_y = 0.2,
                    segment.curvature = -0.1,
                    segment.ncp = 3,
                    # segment.angle = 10,
                    direction = "y", 
                    hjust = "right"
    ) + 
    scale_color_gradientn(colours = c("#3288bd", "#66c2a5","#ffffbf", "#f46d43", "#9e0142"),
                          values = seq(0, 1, 0.2)) +
    scale_fill_gradientn(colours = c("#3288bd", "#66c2a5","#ffffbf", "#f46d43", "#9e0142"),
                         values = seq(0, 1, 0.2)) +
    ggtitle(title, subtitle=subtitle) + xlab("log2 FC") + ylab("-log10 q value") +
    labs(color = 'log2 FC', size = '-log10 q') +
    theme(panel.grid = element_blank(),
          legend.background = element_roundrect(color = "#808080", linetype = 1),
          axis.text = element_text(size = 13, color = "#000000"),
          axis.title = element_text(size = 15),
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) + 
    geom_vline(xintercept=minLFC, linetype=2) + 
    geom_vline(xintercept=-minLFC, linetype=2) + 
    geom_hline(yintercept=minFDR, linetype=4) +
    coord_cartesian(clip = "off") + 
    annotation_custom(
      grob = grid::segmentsGrob(
        y0 = unit(-10, "pt"),
        y1 = unit(-10, "pt"),
        arrow = arrow(angle = 45, length = unit(.2, "cm"), ends = "first"),
        gp = grid::gpar(lwd = 3, col = "#74add1")
      ), 
      xmin = leftarrowlim[1], 
      xmax = leftarrowlim[2],
      ymin = leftarrowlim[3],
      ymax = leftarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::textGrob(
        label = "Down",
        gp = grid::gpar(col = "#74add1")
      ),
      xmin = leftarrowlim[1], 
      xmax = leftarrowlim[2],
      ymin = leftarrowlim[3],
      ymax = leftarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::segmentsGrob(
        y0 = unit(-10, "pt"),
        y1 = unit(-10, "pt"),
        arrow = arrow(angle = 45, length = unit(.2, "cm"), ends = "last"),
        gp = grid::gpar(lwd = 3, col = "#d73027")
      ), 
      xmin = rightarrowlim[1], 
      xmax = rightarrowlim[2],
      ymin = rightarrowlim[3],
      ymax = rightarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::textGrob(
        label = "Up",
        gp = grid::gpar(col = "#d73027")
      ),
      xmin = rightarrowlim[1], 
      xmax = rightarrowlim[2],
      ymin = rightarrowlim[3],
      ymax = rightarrowlim[4]
    ) 
  if (length(axislim) == 4) {
    p1 = p1 + xlim(axislim[1], axislim[2]) + ylim(axislim[3], axislim[4])
  }
  return(p1)
}

# calculate functional scores
get_functional_scores = function(so,
                                 GEXmatrix,
                                 fl = NULL,
                                 module=NULL){
  # select genes corresponding to the query module
  category = fl[[module]]
  # extract normalized gene expression data
  data = GEXmatrix
  # find existed markers under the query module
  category.shared = intersect(category,rownames(data))
  # find markers not existed in the dataset under the module
  category.missing = setdiff(category,rownames(data))
  # extract normalized expression of genes of interest
  df <-
    data %>%
    t() %>%
    as.data.frame() %>%
    select(category.shared)
  # get averages for genes of interest
  df$avg_genes = rowMeans(df)
  if (all(rownames(so@meta.data) == rownames(df))){
    so@meta.data[[paste0('functional_score_of_',module)]] = df$avg_genes
    return(list(so,category.shared,category.missing,length(category.shared)/length(category)))
  } else {
    return('cell barcodes do not match')
  }
}

```

## Load seurat object


```{r load pbmc, eval=T}
pbmc = readRDS('srt.obj.merged.clusterCuratedRefined.classifiedSet.myeloidInfoAdded.metadataIdentifierAdded.FuncSAdded.RDS')
pbmc.copy1 <- pbmc
#pbmc.copy1 -> pbmc
pbmc.metadata = pbmc@meta.data
pbmc.mld2 = readRDS('pbmc.mld2.SexAgeAdded.FuncSAdded.RDS')
pbmc.mld2.copy1 <- pbmc.mld2
#pbmc.mld2.copy1 -> pbmc.mld2
```


# UMAP


## Full UMAP (all cells; no subsampling)


```{r umap colors}
# Define a custom color palette
# Replace these colors with the colors you want for each cluster
custom_colors_fullUmap <- c(
  "CD14 Mono"     = "#E41A1C",  # Red
  "CD16 Mono"     = "#377EB8",  # Blue
  "cDC"           = "#4DAF4A",  # Green
  "pDC"           = "#984EA3",  # Purple
  "B Naive"       = "#FF7F00",  # Orange
  "B memory"      = "#FFFF33",  # Yellow
  "Plasmablast B" = "#A65628",  # Brown
  "CD4 T Naive"   = "#F781BF",  # Pink
  "CD4 T Memory"  = "#999999",  # Grey
  "CD8 T Naive"   = "#A6CEE3",  # Light blue
  "CD8 T Memory"  = "#B2DF8A",  # Light green
  "MAIT"          = "#FB9A99",  # Light red
  "NK"            = "#CAB2D6",  # Light purple
  "gdT"           = "#FFD92F"   # Light yellow
)

custom_colors_mld <- c(
  "0" = "#33A02C",  # Dark green
  "1" = "#6A3D9A",  # Dark purple
  "2" = "#FF6600",  # Bright orange
  "3" = "#1F78B4",  # Dark blue
  "4" = "#B15928",  # Rust
  "5" = "#E31A1C",  # Bright red
  "6" = "#FFCC00",  # Bright yellow
  "7" = "#B2DF8A"   # Pale green
)

```

```{r umap full umap no subsampling,fig.width=6,fig.height=6}
DimPlot(pbmc, reduction = 'wnn.umap', group.by = 'group', label = F, label.size = 5, repel = T)
DimPlot(subset(pbmc, group == 'trisomy8' & trisomy %in% c('NORMAL','TRISOMY')),reduction = 'wnn.umap', group.by = 'trisomy', label = F, label.size = 5, repel = T) + ggtitle('Trisomy vs Disomy cells in Trisomy 8 patients')
DimPlot(pbmc, reduction = 'wnn.umap', group.by = 'curated_clusters4', label = T, label.size = 3, repel = T,
        cols = custom_colors_fullUmap)
```

## Full UMAP (with subsampling to match cell numbers between Trisomy 8 and HC)

```{r full umap with subsampling}
pbmc.subsampled = pbmc
Idents(pbmc.subsampled) <- "group"
table(pbmc.subsampled@meta.data$group)
cat("subsample each group to 26937 cells")
pbmc.subsampled <- subset(pbmc.subsampled, downsample = 26937)
table(pbmc.subsampled@meta.data$group)
DimPlot(pbmc.subsampled, reduction = 'wnn.umap', group.by = 'group', label = F, label.size = 5, repel = T)
DimPlot(subset(pbmc.subsampled, group == 'trisomy8' & trisomy %in% c('NORMAL','TRISOMY')),reduction = 'wnn.umap', group.by = 'trisomy', label = F, label.size = 5, repel = T) + ggtitle('Trisomy vs Disomy cells in Trisomy 8 patients')
DimPlot(pbmc.subsampled, reduction = 'wnn.umap', group.by = 'curated_clusters4', label = T, label.size = 3, repel = T,
        cols = custom_colors_fullUmap)
```
```{r full umap with subsampling and split,fig.width=10,fig.height=6}
DimPlot(pbmc.subsampled, reduction = 'wnn.umap', group.by = 'curated_clusters4', split.by = 'group',label = T, label.size = 3, repel = T,cols = custom_colors_fullUmap)
```

## Myeloid cells (without subsampling)

```{r umap mld umap no subsampling,fig.width=9,fig.height=6}
DimPlot(pbmc.mld2, reduction = "wnn.umap.myeloid", group.by = 'group', label = F, label.size = 5, repel = T)
DimPlot(pbmc.mld2, reduction = "wnn.umap.myeloid", group.by = 'myeloid_Clusters', label = T, label.size = 3, repel = T, cols = custom_colors_mld)
DimPlot(pbmc.mld2, reduction = 'wnn.umap.myeloid', group.by = 'myeloid_Clusters', split.by = 'group', label = F, repel = T, cols = custom_colors_mld)
```


## Myeloid cells(with subsampling to match cell numbers between Trisomy 8 and HC)

```{r mld umap with subsampling}
pbmc.mld2.subsampled = pbmc.mld2
Idents(pbmc.mld2.subsampled) <- "group"
table(pbmc.mld2.subsampled@meta.data$group)
cat("subsample each group to 4775 cells")
pbmc.mld2.subsampled <- subset(pbmc.mld2.subsampled, downsample = 4775)
table(pbmc.mld2.subsampled@meta.data$group)
DimPlot(pbmc.mld2.subsampled, reduction = 'wnn.umap.myeloid', group.by = 'group', label = F, label.size = 5, repel = T)
DimPlot(pbmc.mld2.subsampled, reduction = 'wnn.umap.myeloid', group.by = 'myeloid_Clusters', label = F, repel = T, cols = custom_colors_mld)
DimPlot(pbmc.mld2.subsampled, reduction = 'wnn.umap.myeloid', group.by = 'myeloid_Clusters', split.by = 'group', label = F, repel = T, cols = custom_colors_mld)
```


# Cluster Frequency

Bar plot with the average frequency of each cluster in TRIAD (one bar) and control (one bar) for the full PBMC clustering and myeloid clustering

## All cells, by subject

```{r freq cal}
cluster.freq = cluster_freq(pbmc,"orig.ident",cluster = 'curated_clusters4')
```

```{r freq table}
create_dt(cluster.freq, caption = "Cluster frequency for all subjects (using curated_clusters4)")
```

```{r freq plot,fig.width=6,fig.height=6}
ggplot(data = cluster.freq, mapping = aes(x = orig.ident, y = `relative_frequency`,fill = curated_clusters4)) +
  geom_bar(stat="identity",color = 'black') +
  scale_fill_manual(values = custom_colors_fullUmap) +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab(" ") + ylab('Cluster frequency') +
  guides(fill=guide_legend(nrow=4))
```

## All cells, by group

```{r freq by group}
cluster.freq_byGroup <-
  pbmc@meta.data %>%
  group_by(group, curated_clusters4) %>%
  summarise(count=n()) %>%
  mutate(relative_frequency = count/sum(count))

create_dt(cluster.freq_byGroup, caption = "Cluster frequency by groups (using curated_clusters4)")
```

```{r freq plot by group,fig.width=6,fig.height=6}
ggplot(data = cluster.freq_byGroup, mapping = aes(x = group, y = `relative_frequency`,fill = curated_clusters4)) +
  geom_bar(stat="identity",color = 'black') +
  scale_fill_manual(values = custom_colors_fullUmap) +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab(" ") + ylab('Cluster frequency') +
  guides(fill=guide_legend(nrow=4))

```

## Myeloid cells, by subject 

myeloid cells only, based on myeloid_Clusters column.

```{r m freq}
cluster.freq.myeloids = cluster_freq(pbmc.mld2,'orig.ident','myeloid_Clusters')
```

```{r myl freq table}
create_dt(cluster.freq.myeloids, caption = "Myeloid cell cluster frequency by subjects (using myeloid_Clusters column)")
```

```{r myl fre plot,fig.width=6,fig.height=6}
ggplot(data = cluster.freq.myeloids, mapping = aes(x = orig.ident, y = `relative_frequency`,fill = myeloid_Clusters)) +
  geom_bar(stat="identity",color = 'black') +
  scale_fill_manual(values = custom_colors_mld) +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab("Subject") + ylab('Cluster frequency') +
  guides(fill=guide_legend(nrow=2))
```


## Myeloid cells, by group

```{r freq by group mld}
cluster.freq.myeloids_byGroup <-
  pbmc.mld2@meta.data %>%
  group_by(group, myeloid_Clusters) %>%
  summarise(count=n()) %>%
  mutate(relative_frequency = count/sum(count))

create_dt(cluster.freq.myeloids_byGroup, caption = "Myeloids cells subclusters frequency by groups (using myeloid_Clusters)")
```


```{r freq plot by group mld,fig.width=5,fig.height=6}
ggplot(data = cluster.freq.myeloids_byGroup, mapping = aes(x = group, y = `relative_frequency`,fill = myeloid_Clusters)) +
  geom_bar(stat="identity",color = 'black') +
  scale_fill_manual(values = custom_colors_mld) +
  theme_classic() +
  theme(legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
    xlab(" ") + ylab('Cluster frequency') +  guides(fill=guide_legend(nrow=2,title = 'Myeloid cells subclusters'))

```

# Key genes and ADT markers for cell cluster annotation

## Genes

** Selected genes **

"CD14", # CD14 Monocytes

"CDKN1C", # CD16 Monocytes

"FCER1A", # cDC

"SERPINF1", # pDC

"CD79A", "CD79B", # B

"MZB1", # Plasmablast B

'CD4','IL7R', 'MAL', # CD4T

"CCR7","CD27", # CD4T Naive/Memory

"CD8B", "CD8A", # CD8T

"NKG7", "KLRD1", "TYROBP", #NK

"NCR3","GZMK", # MAIT

"TRDC","TRDV2" # dgT

  
  
```{r key genes,fig.width=10,fig.height=6}
keyGenes = c(
  "CD14", # CD14 Monocytes
  "CDKN1C", # CD16 Monocytes
  "FCER1A", # cDC
  "SERPINF1", # pDC
  "CD79A", "CD79B", # B
  "MZB1", # Plasmablast B
  'CD4','IL7R', 'MAL', # CD4T
  "CCR7","CD27", # CD4T Naive/Memory
  "CD8B", "CD8A", # CD8T
  "NKG7", "KLRD1", "TYROBP", #NK
  "NCR3","GZMK", # MAIT
  "TRDC","TRDV2" # dgT
  )
cluster_order = c("CD14 Mono",
                  "CD16 Mono",
                  'cDC',
                  'pDC',
                  "B Naive",
                  "B memory",
                  "Plasmablast B",
                  "CD4 T Naive",
                  "CD4 T Memory",
                  "CD8 T Naive",
                  "CD8 T Memory",
                  'MAIT',
                  "NK",
                  'gdT'
                  )
jjDotPlot(object = pbmc,
          gene = keyGenes,
          gene.order = keyGenes,
          cluster.order = rev(cluster_order),
          id = 'curated_clusters4',ytree = F)

```

## ADTs

** Selected ADT markers **

"CD14", # CD14 Monocytes

"CD16", # CD16 Monocytes

"CD19","IgM", # B

"CD27","IgD", # B Naive/Memory

'CD4', #CD4T

"CD45RA",  # CD4T Naive/Memory

'CD8', #CD8T

"CD3" # NK
  
```{r key ADTs,fig.width=10,fig.height=6}
keyADTs = c(
  "CD14", # CD14 Monocytes
  "CD16", # CD16 Monocytes
  "CD19","IgM", # B
  "CD27","IgD", # B Naive/Memory
  'CD4', #CD4T
  "CD45RA",  # CD4T Naive/Memory
  'CD8', #CD8T
  "CD3" # NK
)

jjDotPlot(object = pbmc,assay = 'ADT',
          gene = keyADTs,
          id = 'curated_clusters4',
          gene.order = keyADTs,
          cluster.order = rev(cluster_order),
          dot.col = c('white','blue'),
          ytree = F)

```

# Functional scores for the selected pathways

```{r fs pw selection}
# (this step takes a long time, so do it once for all following iterations)
geneSet.kegg = gmtPathways('c2.cp.kegg_medicus.v2023.2.Hs.symbols.gmt')
geneSet.reactome =gmtPathways('c2.cp.reactome.v2023.2.Hs.symbols.gmt') 
geneSet.go = gmtPathways('c5.go.bp.v2023.2.Hs.symbols.gmt')
geneSet.hallmark = gmtPathways('h.all.v2023.2.Hs.symbols.gmt')
geneSet.merged4 = c(geneSet.kegg,geneSet.reactome,geneSet.go,geneSet.hallmark)

pw = c('HALLMARK_INTERFERON_GAMMA_RESPONSE',
'HALLMARK_Interferon alpha response',
'GOBP_Response to type II INTERFERON',
'GOBP_Response to type I INTERFERON',
'HALLMARK_TNFA signaling via NFKB',
'HALLMARK_IL6_JAK_STAT3_Signaling',
'HALLMARK_Apoptosis',
"GOBP_RESPONSE_TO_TUMOR_NECROSIS_FACTOR",
'GOBP_Monocyte chemotaxis',
'REACTOME_Translation',
  "GOBP_RESPONSE_TO_INTERLEUKIN_1",
  "GOBP_POSITIVE_REGULATION_OF_INTERLEUKIN_1_BETA_PRODUCTION",
  "GOBP_INTERLEUKIN_1_MEDIATED_SIGNALING_PATHWAY",
  "GOBP_INTERLEUKIN_1_PRODUCTION",
  "GOBP_CELLULAR_RESPONSE_TO_INTERLEUKIN_1",
  "REACTOME_INTERLEUKIN_1_SIGNALING"
)

# Capitalize all letters and replace spaces with underscores
pw <- toupper(gsub(" ", "_", pw))
# check if everything matched
# setdiff(pw,names(geneSet.merged4))
```

## Myeloid subset

```{r fs plot prep mld}

fs.raw = pbmc.mld2@meta.data
cat(paste0('number of cells in functional score calculation: ',nrow(fs.raw)))
fs.genes = readRDS('functional_score_genes_summary.mld2.session30.rds')

fs.raw = fs.raw %>%
  select(c(group,orig.ident,paste0('functional_score_of_',pw)))
names(fs.raw) = gsub('functional_score_of_','',names(fs.raw))

# Calculate the averages for each group and orig.ident
# by group
fs.group.avg <- fs.raw %>%
  group_by(group) %>%
  summarise(
    HALLMARK_INTERFERON_GAMMA_RESPONSE = mean(HALLMARK_INTERFERON_GAMMA_RESPONSE, na.rm = TRUE),
    HALLMARK_INTERFERON_ALPHA_RESPONSE = mean(HALLMARK_INTERFERON_ALPHA_RESPONSE, na.rm = TRUE),
    GOBP_RESPONSE_TO_TYPE_II_INTERFERON = mean(GOBP_RESPONSE_TO_TYPE_II_INTERFERON, na.rm = TRUE),
    GOBP_RESPONSE_TO_TYPE_I_INTERFERON = mean(GOBP_RESPONSE_TO_TYPE_I_INTERFERON, na.rm = TRUE),
    HALLMARK_TNFA_SIGNALING_VIA_NFKB = mean(HALLMARK_TNFA_SIGNALING_VIA_NFKB, na.rm = TRUE),
    HALLMARK_IL6_JAK_STAT3_SIGNALING = mean(HALLMARK_IL6_JAK_STAT3_SIGNALING, na.rm = TRUE),
    HALLMARK_APOPTOSIS = mean(HALLMARK_APOPTOSIS, na.rm = TRUE),
    GOBP_RESPONSE_TO_TUMOR_NECROSIS_FACTOR = mean(GOBP_RESPONSE_TO_TUMOR_NECROSIS_FACTOR, na.rm = TRUE),
    GOBP_MONOCYTE_CHEMOTAXIS = mean(GOBP_MONOCYTE_CHEMOTAXIS, na.rm = TRUE),
    REACTOME_TRANSLATION = mean(REACTOME_TRANSLATION, na.rm = TRUE),
    GOBP_RESPONSE_TO_INTERLEUKIN_1 = mean(GOBP_RESPONSE_TO_INTERLEUKIN_1, na.rm = TRUE),
    GOBP_POSITIVE_REGULATION_OF_INTERLEUKIN_1_BETA_PRODUCTION = mean(GOBP_POSITIVE_REGULATION_OF_INTERLEUKIN_1_BETA_PRODUCTION, na.rm = TRUE),
    GOBP_INTERLEUKIN_1_MEDIATED_SIGNALING_PATHWAY = mean(GOBP_INTERLEUKIN_1_MEDIATED_SIGNALING_PATHWAY, na.rm = TRUE),
    GOBP_INTERLEUKIN_1_PRODUCTION = mean(GOBP_INTERLEUKIN_1_PRODUCTION, na.rm = TRUE),
    GOBP_CELLULAR_RESPONSE_TO_INTERLEUKIN_1 = mean(GOBP_CELLULAR_RESPONSE_TO_INTERLEUKIN_1, na.rm = TRUE),
    REACTOME_INTERLEUKIN_1_SIGNALING = mean(REACTOME_INTERLEUKIN_1_SIGNALING, na.rm = TRUE)
  ) 
# by subject
fs.sample.avg <- fs.raw %>%
  group_by(orig.ident) %>%
  summarise(
    HALLMARK_INTERFERON_GAMMA_RESPONSE = mean(HALLMARK_INTERFERON_GAMMA_RESPONSE, na.rm = TRUE),
    HALLMARK_INTERFERON_ALPHA_RESPONSE = mean(HALLMARK_INTERFERON_ALPHA_RESPONSE, na.rm = TRUE),
    GOBP_RESPONSE_TO_TYPE_II_INTERFERON = mean(GOBP_RESPONSE_TO_TYPE_II_INTERFERON, na.rm = TRUE),
    GOBP_RESPONSE_TO_TYPE_I_INTERFERON = mean(GOBP_RESPONSE_TO_TYPE_I_INTERFERON, na.rm = TRUE),
    HALLMARK_TNFA_SIGNALING_VIA_NFKB = mean(HALLMARK_TNFA_SIGNALING_VIA_NFKB, na.rm = TRUE),
    HALLMARK_IL6_JAK_STAT3_SIGNALING = mean(HALLMARK_IL6_JAK_STAT3_SIGNALING, na.rm = TRUE),
    HALLMARK_APOPTOSIS = mean(HALLMARK_APOPTOSIS, na.rm = TRUE),
    GOBP_RESPONSE_TO_TUMOR_NECROSIS_FACTOR = mean(GOBP_RESPONSE_TO_TUMOR_NECROSIS_FACTOR, na.rm = TRUE),
    GOBP_MONOCYTE_CHEMOTAXIS = mean(GOBP_MONOCYTE_CHEMOTAXIS, na.rm = TRUE),
    REACTOME_TRANSLATION = mean(REACTOME_TRANSLATION, na.rm = TRUE),
    GOBP_RESPONSE_TO_INTERLEUKIN_1 = mean(GOBP_RESPONSE_TO_INTERLEUKIN_1, na.rm = TRUE),
    GOBP_POSITIVE_REGULATION_OF_INTERLEUKIN_1_BETA_PRODUCTION = mean(GOBP_POSITIVE_REGULATION_OF_INTERLEUKIN_1_BETA_PRODUCTION, na.rm = TRUE),
    GOBP_INTERLEUKIN_1_MEDIATED_SIGNALING_PATHWAY = mean(GOBP_INTERLEUKIN_1_MEDIATED_SIGNALING_PATHWAY, na.rm = TRUE),
    GOBP_INTERLEUKIN_1_PRODUCTION = mean(GOBP_INTERLEUKIN_1_PRODUCTION, na.rm = TRUE),
    GOBP_CELLULAR_RESPONSE_TO_INTERLEUKIN_1 = mean(GOBP_CELLULAR_RESPONSE_TO_INTERLEUKIN_1, na.rm = TRUE),
    REACTOME_INTERLEUKIN_1_SIGNALING = mean(REACTOME_INTERLEUKIN_1_SIGNALING, na.rm = TRUE)
  ) 
```

### Functional scores and gene recruitment rate

```{r fs show tables mld}
create_dt(fs.genes,'Module score gene recruitment rates')
create_dt(fs.sample.avg,'Module scores averages by sample')
create_dt(fs.group.avg,'Module scores averages by group')
```



### Heatmaps

The functional scores were standardized (scale x to zero mean and unit variance) by row (by function) before plotting.


```{r fs heatmap mld,fig.width=15,fig.height=6}
fs.sample.avg.scaled <- fs.sample.avg %>%
  as.data.frame() %>%
  setrowname() %>%
  t() %>%
  as.data.frame() %>%
  decostand(.,MARGIN = 1,method = 'standardize')
  
min(fs.sample.avg.scaled)
max(fs.sample.avg.scaled)
col_fun = colorRamp2(c(-1.5, 0, 2.5), c("#2B5C8AFF", "#FFF6EDFF", "#AC4522FF"))

fs.sample.avg.scaled.control = as.matrix(select(fs.sample.avg.scaled, c('HV-1', 'HV-2', 'HV-3')))
fs.sample.avg.scaled.t8 = as.matrix(select(fs.sample.avg.scaled, -c('HV-1', 'HV-2', 'HV-3')))

fs.sample.avg.scaled.control.plt = Heatmap(fs.sample.avg.scaled.control, name = 'Scaled module score',
                                    column_title = "Control",
                                    show_row_names = T,show_row_dend = F,
                                    show_column_names = T, show_column_dend = F,
                                    cluster_columns = F,cluster_rows = F,
                                    col = col_fun,
                                    border_gp = gpar(col = "black", lty = 1),
                                    row_names_side = "left",
                                    row_names_max_width = unit(20, "cm"))
fs.sample.avg.scaled.t8.plt = Heatmap(fs.sample.avg.scaled.t8, name = 'Scaled module score',
                               column_title = "Trisomy 8",
                               show_row_names = T,show_row_dend = F,
                               show_column_names = T, show_column_dend = F,
                               cluster_columns = F,cluster_rows = F,
                               col = col_fun,
                               border_gp = gpar(col = "black", lty = 1),
                              row_names_side = "left",
                               row_names_max_width = unit(20, "cm"))
fs.plt = fs.sample.avg.scaled.control.plt +  fs.sample.avg.scaled.t8.plt
draw(fs.plt,merge_legend = TRUE, heatmap_legend_side = "right" )

# pdf('session30.functionalScore.heatmap.pdf', width =10, height = 5)
# par(mar = c(5,5,5,5) + 0.1)
# draw(fs.plt,merge_legend = TRUE, heatmap_legend_side = "right" )
# dev.off()

create_dt(fs.sample.avg.scaled,'Module score (scaled by row)')
```


### Feature plot (all samples in one UMAP)

```{r fs feature mld plots all in one,results='asis',fig.width=4,fig.height=4}
for (pathway in pw){
  cat('\n')
  cat('####', pathway, '\n')
  cat('\n')
  print(FeaturePlot(pbmc.mld2, features = paste0('functional_score_of_',pathway),reduction = 'wnn.umap.myeloid')& theme(plot.title = element_text(size = 5, face = "bold")))
  cat('\n')
}
```
