---
title: "VDJ T cell analysis (scRepertoire)"
date: "`r Sys.Date()`"
author: Yue (Gary) Zhang, PhD
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = T, message = FALSE) 
```

# Preparation

## Load packages and set variables

```{r Initiate, results='hide'}
## initiate ---------------------------------------------------------
# use the following line to set the library path to /data/zhangy68/R/4.3;
# this path is writtable and used by R in sinteractive mode
# this line is nessacery if running interactive web-based Rstudio
.libPaths('/data/zhangy68/R/4.3')

packages_to_load = c(
  "dplyr",
  "Seurat",
  "patchwork",
  "ggplot2",
  "ggrepel",
  "readxl",
  "future",
  "cowplot",
  "Azimuth",
  "dsb",
  "corrplot",
  "DT",
  "tidyr",
  "ggpubr",
  "scRNAtoolVis",
  "purrr",
  "limma",
  "wesanderson",
  "reshape",
  "edgeR",
  'ggfun',
  'ggrepel',
  'fgsea',
  'circlize',
  'ComplexHeatmap',
  'vegan',
  "scRepertoire"
)
# Load packages without printing output
invisible(suppressPackageStartupMessages({
  lapply(packages_to_load, library, character.only = TRUE)
}))

# set variables 
setwd('./')
set.seed('123')
col21 <- rev(c("tomato1","darkblue","turquoise1","lightblue","darkred","mediumblue","purple","bisque",
               "greenyellow","yellow","violetred2","darkgreen","darkgoldenrod1","deeppink3","cadetblue4",
               "orchid2","seagreen3","purple4","dodgerblue2","red","gray27"))

```

## Customized functions

```{r function section}
# create a function to generate cluster frequency tables
cluster_freq = function(pbmc,orig.ident,cluster,sex=F){
  freq <- pbmc@meta.data %>%
    group_by(!!sym(orig.ident), !!sym(cluster)) %>%
    summarise(count=n()) %>%
    mutate(relative_frequency = count/sum(count))
  freq$group = ifelse(freq$orig.ident %in% c('HV-1', 'HV-2', 'HV-3'), 'control', 'trisomy8')
  if(sex == T){
    freq$sex = cov$sex[match(freq$orig.ident,cov$sample)]
    freq$group_sex = paste0(freq$group,'_',freq$sex)
    freq$group_sex = factor(freq$group_sex, levels = c('control_F','trisomy8_F','control_M','trisomy8_M'))
  }
  freq$orig.ident = factor(freq$orig.ident, levels = c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4'))
  return(freq)
}
# custom function to plot DE genes between Trisomy and Normal by clusters
plot_de_volcano_scatter = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Trisomy vs Normal') {
  # remove everything after "..." in gene names; bind_rows automatically added these to avoid duplicate rownames.
  df$genes = sub(pattern = '\\.\\.\\.*', replacement = '',x = rownames(df))
  # calculate log10 q value. cap log10 q value at 5 (which is 0.00001 in q val)
  df$p_val_adj_log10 = -log10(df$p_val_adj)
  df$p_val_adj_log10_capped = ifelse(df$p_val_adj_log10 < 5, df$p_val_adj_log10, 5)
  # generate summary statistics to show on caption
  minFDR <- -log10(maxq)
  sigs <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) >= minLFC
  sigLowFC <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) < minLFC
  
  cnt_plus <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC >= minLFC)
  cnt_neg <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC <= -minLFC)
  if(showsub) {
    subtitle <- paste0(sum(sigs), " DE genes: ", cnt_plus, " Pos and ", cnt_neg, " Neg",
                       " (FDR<", maxq, ", log2FC>", minLFC, ")")
  } else {
    subtitle <- NULL
  }
  
  p1 = ggplot(df) +
    geom_point(aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5) + 
    geom_point(data=df[sigs,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigcol) +
    geom_point(data=df[sigLowFC,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigLowFCcol) +
    ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("log10 FDR") +
    facet_wrap(.~Celltype_tested,nrow=round(length(unique(df$Celltype_tested))/2))+
    theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) + 
    geom_vline(xintercept=minLFC, col=sigcol, linetype="dashed", size=0.5) + 
    geom_vline(xintercept=-minLFC, col=sigcol, linetype="dashed", size=0.5) + 
    geom_hline(yintercept=minFDR, col=sigcol, linetype="dashed", size=0.5)
  
  df.s = subset(df, p_val_adj_log10 >minFDR & abs(avg_log2FC) > minLFC)
  df.s = df.s[order(df.s$avg_log2FC, decreasing = TRUE), ]
  df.s$genes = factor(df.s$genes,levels = unique(df.s$genes))
  p2 = ggplot(df.s,mapping = aes(x=genes,y=avg_log2FC,color=p_val_adj_log10_capped)) +
    geom_point() +
    facet_wrap(~ Celltype_tested, scales = 'free') +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    coord_flip() +
    ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("genes")
  return(list(p1,p2))
}

create_dt <- function(df,caption){
  datatable(df,
            caption = caption,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All")),
                           initComplete = JS(
    "function(settings, json) {",
    "$('caption').css({'font-size': '20px', 'font-family': 'Arial', 'font-weight': 'bold'});",
    "}")))
}

# function to set a column as rownames and remove that column from the df; default is setting the first column as the rowname
setrowname =  function(df,col = 1){
  if (is.numeric(col)) {
    # print('it is a number')
    rownames(df) = df[,col]
    df = df[,-col]
  } else {
    # print('it is a string')
    rownames(df) = df[,which(names(df) %in% c(col))]
    df = df[,-which(names(df) %in% c(col))]
  }
  return(df)
} 

# function to get colors from a palatte
getcol = function(group,pal = "Darjeeling1",shift_col_palatte=0){
  groupcol = as.numeric(factor(group))
  groupcol = wes_palette(pal)[groupcol+shift_col_palatte]
  return(groupcol)
}


# Volcano plot
plot_de_volcano_scatter3 = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Volcano plot',
                                    axislim,leftarrowlim,rightarrowlim) {
  # calculate log10 q value. 
  df$qval_log10 = -log10(df$`adj.P.Val`)
  # generate summary statistics to show on caption
  minFDR <- -log10(maxq)
  sigs <- df$qval_log10 >= minFDR & abs(df$logFC) >= minLFC
  sigLowFC <- df$qval_log10 >= minFDR & abs(df$logFC) < minLFC
  
  cnt_plus <- sum(df$qval_log10 >= minFDR & df$logFC >= minLFC)
  cnt_neg <- sum(df$qval_log10 >= minFDR & df$logFC <= -minLFC)
  if(showsub) {
    subtitle <- paste0(sum(sigs), " DEGs: ", cnt_plus, " + and ", cnt_neg, " -",
                       " (q<", maxq, ", log2FC>", minLFC, ")")
  } else {
    subtitle <- NULL
  }
  
  p1 = ggplot(df) +
    geom_point(aes(x=logFC, y=qval_log10, color = logFC, size = qval_log10)) + 
    geom_point(data =  df %>%
                 subset(qval_log10 >= minFDR) %>%
                 dplyr::arrange(desc(logFC)) %>%
                 dplyr::slice(1:5),
               aes(x = logFC, y = qval_log10,
                   # fill = log2FoldChange,
                   size = qval_log10),
               shape = 21, show.legend = F, color = "#000000") +
    geom_text_repel(data =  df %>%
                      subset(qval_log10 >= minFDR) %>%
                      dplyr::arrange(desc(logFC)) %>%
                      dplyr::slice(1:5),
                    aes(x = logFC, y = qval_log10,label = geneID),
                    box.padding = 0.5,
                    nudge_x = 0.5,
                    nudge_y = 0.2,
                    segment.curvature = -0.1,
                    segment.ncp = 3,
                    # segment.angle = 10,
                    direction = "y", 
                    hjust = "left"
    ) + 
    geom_point(data =  df %>%
                 subset(qval_log10 >= minFDR) %>%
                 dplyr::arrange(logFC) %>%
                 dplyr::slice(1:5),
               aes(x = logFC, y = qval_log10,
                   # fill = log2FoldChange,
                   size = qval_log10),
               shape = 21, show.legend = F, color = "#000000") +
    geom_text_repel(data =  df %>%
                      subset(qval_log10 >= minFDR) %>%
                      dplyr::arrange(logFC) %>%
                      dplyr::slice(1:5),
                    aes(x = logFC, y = qval_log10,label = geneID),
                    box.padding = 0.5,
                    nudge_x = 0.5,
                    nudge_y = 0.2,
                    segment.curvature = -0.1,
                    segment.ncp = 3,
                    # segment.angle = 10,
                    direction = "y", 
                    hjust = "right"
    ) + 
    scale_color_gradientn(colours = c("#3288bd", "#66c2a5","#ffffbf", "#f46d43", "#9e0142"),
                          values = seq(0, 1, 0.2)) +
    scale_fill_gradientn(colours = c("#3288bd", "#66c2a5","#ffffbf", "#f46d43", "#9e0142"),
                         values = seq(0, 1, 0.2)) +
    ggtitle(title, subtitle=subtitle) + xlab("log2 FC") + ylab("-log10 q value") +
    labs(color = 'log2 FC', size = '-log10 q') +
    theme(panel.grid = element_blank(),
          legend.background = element_roundrect(color = "#808080", linetype = 1),
          axis.text = element_text(size = 13, color = "#000000"),
          axis.title = element_text(size = 15),
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) + 
    geom_vline(xintercept=minLFC, linetype=2) + 
    geom_vline(xintercept=-minLFC, linetype=2) + 
    geom_hline(yintercept=minFDR, linetype=4) +
    coord_cartesian(clip = "off") + 
    annotation_custom(
      grob = grid::segmentsGrob(
        y0 = unit(-10, "pt"),
        y1 = unit(-10, "pt"),
        arrow = arrow(angle = 45, length = unit(.2, "cm"), ends = "first"),
        gp = grid::gpar(lwd = 3, col = "#74add1")
      ), 
      xmin = leftarrowlim[1], 
      xmax = leftarrowlim[2],
      ymin = leftarrowlim[3],
      ymax = leftarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::textGrob(
        label = "Down",
        gp = grid::gpar(col = "#74add1")
      ),
      xmin = leftarrowlim[1], 
      xmax = leftarrowlim[2],
      ymin = leftarrowlim[3],
      ymax = leftarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::segmentsGrob(
        y0 = unit(-10, "pt"),
        y1 = unit(-10, "pt"),
        arrow = arrow(angle = 45, length = unit(.2, "cm"), ends = "last"),
        gp = grid::gpar(lwd = 3, col = "#d73027")
      ), 
      xmin = rightarrowlim[1], 
      xmax = rightarrowlim[2],
      ymin = rightarrowlim[3],
      ymax = rightarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::textGrob(
        label = "Up",
        gp = grid::gpar(col = "#d73027")
      ),
      xmin = rightarrowlim[1], 
      xmax = rightarrowlim[2],
      ymin = rightarrowlim[3],
      ymax = rightarrowlim[4]
    ) 
  if (length(axislim) == 4) {
    p1 = p1 + xlim(axislim[1], axislim[2]) + ylim(axislim[3], axislim[4])
  }
  return(p1)
}
```

## Load seurat object

```{r load pbmc, eval=T}
pbmc = readRDS('srt.obj.merged.clusterCuratedRefined.classifiedSet.myeloidInfoAdded.metadataIdentifierAdded.FuncSAdded.RDS')
pbmc.copy1 <- pbmc
#pbmc.copy1 -> pbmc

# pbmc.mld2 = readRDS('pbmc.mld2.SexAgeAdded.RDS')
# pbmc.mld2.copy1 <- pbmc.mld2
#pbmc.mld2.copy1 -> pbmc.mld2
```

## Loading and Processing Contig Data

```{r load contig}
# vdj t
file_path = '/data/NCBR/projects/NCBR-323/multi-3rd-pass_finalMultiUsingHTODemuxAssignment/download_for_downstream_analysis/VDJ_T_files/filtered_contig_annotations/'
#list.files(file_path)

contig_list = list()

for (i in c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4')){
  contig_list[[i]] = read.csv(paste0(file_path,i,'.csv'))
}


saveRDS(contig_list,'contig_list.T.session28_2.screpertoire.rds')

```

```{r combine contig list}
group = ifelse(c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4') %in% c('HV-1', 'HV-2', 'HV-3'), 'control', 'trisomy8')
combined <- combineTCR(contig_list, 
                        samples = c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4'), 
                        cells ="T-AB")
combined <- addVariable(combined, 
                      name = "group", 
                      variables = group)
clonecall = 'aa'

```


# Combine with seurat object

Here, I subset the seurat object to only T cells (CD4/8 T Naive/Memory, gdT and MAIT), and then merged with the TCR object.

Using 0.01 as cutoff for unexpanded vs expanded

```{r proportion of sharing cell barcodes between TCR and GEX data, eval = T}
tCell_types= c('CD4 T Memory',  'CD4 T Naive',
                                     'CD8 T Memory',  'CD8 T Naive',
                                      'gdT', 'MAIT')

unexpanded_cutoff = 0.001
unexpanded_label = "unexpanded (0 < X <= 0.001)"
expanded_label = "expanded (0.001 < X <= 1)"
# subset seurat obj to only T cells
pbmc.Tcells = subset(pbmc,curated_clusters4 %in% tCell_types)
# merge TCR and seurat obj.
pbmc.TCR.combined <- combineExpression(combined, 
                                   pbmc.Tcells, 
                                   cloneCall=clonecall, 
                                   group.by = "sample", 
                                   proportion = T, 
                                   cloneTypes=c(unexpanded = unexpanded_cutoff, expanded = 1))
# remove the NAs
pbmc.TCR.combined = subset(pbmc.TCR.combined, cloneType %in% c(unexpanded_label,expanded_label))
#table(pbmc.TCR.combined@meta.data$orig.ident)
```

# UMAP showing clonal space homeostasis


```{r umap homeo}
pbmc.TCR.combined@meta.data$cloneType = factor(pbmc.TCR.combined@meta.data$cloneType,
                                               levels =  c(unexpanded_label,expanded_label))

```

```{r umap plot}
DimPlot(pbmc.TCR.combined, reduction = 'wnn.umap', group.by = 'cloneType', label = F, label.size = 5, repel = T)
DimPlot(pbmc.TCR.combined, reduction = 'wnn.umap', group.by = 'curated_clusters4', label = T, label.size = 5, repel = T)
```

```{r umap plot2,fig.width=10,fig.height=6}
DimPlot(pbmc.TCR.combined, reduction = 'wnn.umap', group.by = 'cloneType', split.by = 'group',label = F, label.size = 5, repel = T)
```

```{r umap plot3,fig.width=10,fig.height=12}
DimPlot(pbmc.TCR.combined, reduction = 'wnn.umap', group.by = 'cloneType', split.by = 'orig.ident', ncol = 3, label = F, label.size = 5, repel = T)
```

# Expanded vs. unexpanded cell frequency plot

## Trisomy vs Disomy cells in Trisomy 8 patients CD8 T memory cells

### Compare the trisomy cell percentage in expanded CD8 T memory cells

Here I used the previous definition of expanded/unexpanded, selected only the CD8 T memory cells with trisomy and disomy (represented as "NORMAL") annotation, and calculated the number and frequency of trisomy and disomy cells within each patient expanded/unexpanded CD8 T memory cells.

```{r exp t8 2}
exp8tm.t8.freq.sample <- pbmc.TCR.combined@meta.data %>%
  subset(curated_clusters4_trisomy %in% c("CD8 T Memory-TRISOMY", "CD8 T Memory-NORMAL")) %>%
  group_by(orig.ident,cloneType,trisomy) %>%
  summarise(count=n()) %>%
  mutate(relative_frequency = count/sum(count))


# add missing rows (if count is zero, that row is missing)
exp8tm.t8.freq.sample.missing = NULL
for (sample in c('TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4')){
  for (exp in c(expanded_label,unexpanded_label)){
    for (triso in c('TRISOMY','NORMAL')){
      if (nrow(subset(exp8tm.t8.freq.sample,orig.ident == sample & trisomy == triso & cloneType == exp)) == 1){
        if (exp8tm.t8.freq.sample$relative_frequency[exp8tm.t8.freq.sample$orig.ident == sample & exp8tm.t8.freq.sample$trisomy == triso & exp8tm.t8.freq.sample$cloneType == exp] == 1){
        triso2 = setdiff(c('TRISOMY','NORMAL'),c(triso))
        exp8tm.t8.freq.sample.missing = as.data.frame(rbind(exp8tm.t8.freq.sample.missing,
                                              c(sample,exp,triso2, 0,0)))
        }
      }
      
    }
  }
}

names(exp8tm.t8.freq.sample.missing) = names(exp8tm.t8.freq.sample)
exp8tm.t8.freq.sample.missing$count = as.numeric(exp8tm.t8.freq.sample.missing$count)
exp8tm.t8.freq.sample.missing$relative_frequency = as.numeric(exp8tm.t8.freq.sample.missing$relative_frequency)

exp8tm.t8.freq.sample = bind_rows(exp8tm.t8.freq.sample,exp8tm.t8.freq.sample.missing)

create_dt(exp8tm.t8.freq.sample,caption = 'Trisomy vs disomy clone type frequency in Trisomy8 patient CD8 T Memory cells')


ggplot(data = exp8tm.t8.freq.sample, mapping = aes(x = orig.ident, y = `relative_frequency`,fill = trisomy)) +
  geom_bar(stat="identity",color = 'black') +
  scale_fill_manual(values = wes_palette("Darjeeling2", n = 2)) +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom',
        legend.text = element_text(size = 12),
        legend.key.size = unit(0.3,'cm'),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=10, face = 'bold', color = 'black'),
        axis.text.y = element_text(size=15, face = 'bold', color = 'black'),
        axis.title=element_text(size=18,face="bold"),
        strip.text.x = element_text(size = 15, color = "black", face = "bold"),
        strip.text.y = element_text(size = 15, color = "black", face = "bold"),
        strip.background = element_rect(color="black", fill="grey95", size=1.5, linetype="solid"))+
  xlab("Subject") + ylab('Clone type frequency') +
  guides(fill=guide_legend(nrow=2))+
  facet_wrap(~cloneType)

ggboxplot(subset(exp8tm.t8.freq.sample,trisomy == 'TRISOMY'), 
          x = "cloneType", y = "relative_frequency",
          color = "cloneType", palette = wes_palette("Darjeeling1", n = 2),
          add = "jitter", # Adding jitter to show data points
          ylab = "Frequency", xlab = " ",
          title = "Proportion of trisomy cells in CD8 T memory cells")+
  stat_compare_means(method = "wilcox.test")
```


# Session info and save R workspace

```{r session info}
sessionInfo()
#save.image("session_28_2_TCR_scRep_workspace.RData")
#saveRDS(pbmc.TCR.combined,'srt.obj.merged.clusterCuratedRefined.classifiedSet.myeloidInfoAdded.metadataIdentifierAdded.FuncSAdded.TCRadded.RDS')
```
