---
title: "BCR Immcantation"
date: "`r Sys.Date()`"
author: Yue (Gary) Zhang, PhD
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = T, message = FALSE) 
```

# Preparation

## Load packages and set variables

```{r Initiate, results='hide'}
## initiate ---------------------------------------------------------
# use the following line to set the library path to /data/zhangy68/R/4.3;
# this path is writtable and used by R in sinteractive mode
# this line is nessacery if running interactive web-based Rstudio
.libPaths('/data/zhangy68/R/4.3')

packages_to_load = c(
  "dplyr",
  "Seurat",
  "patchwork",
  "ggplot2",
  "ggrepel",
  "readxl",
  "future",
  "cowplot",
  "Azimuth",
  "dsb",
  "corrplot",
  "DT",
  "tidyr",
  "ggpubr",
  "scRNAtoolVis",
  "purrr",
  "limma",
  "wesanderson",
  "reshape",
  "edgeR",
  'ggfun',
  'ggrepel',
  'fgsea',
  'circlize',
  'ComplexHeatmap',
  'vegan',
  "immunarch"
)
# Load packages without printing output
invisible(suppressPackageStartupMessages({
  lapply(packages_to_load, library, character.only = TRUE)
}))

# Immcantation required packages
suppressPackageStartupMessages(library(airr))
suppressPackageStartupMessages(library(alakazam))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(dowser))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(scoper))
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(shazam))

# set variables 
setwd('./')
set.seed('123')
col21 <- rev(c("tomato1","darkblue","turquoise1","lightblue","darkred","mediumblue","purple","bisque",
               "greenyellow","yellow","violetred2","darkgreen","darkgoldenrod1","deeppink3","cadetblue4",
               "orchid2","seagreen3","purple4","dodgerblue2","red","gray27"))

```

## Customized functions

```{r function section}
# create a function to generate cluster frequency tables
cluster_freq = function(pbmc,orig.ident,cluster,sex=F){
  freq <- pbmc@meta.data %>%
    group_by(!!sym(orig.ident), !!sym(cluster)) %>%
    summarise(count=n()) %>%
    mutate(relative_frequency = count/sum(count))
  freq$group = ifelse(freq$orig.ident %in% c('HV-1', 'HV-2', 'HV-3'), 'control', 'trisomy8')
  if(sex == T){
    freq$sex = cov$sex[match(freq$orig.ident,cov$sample)]
    freq$group_sex = paste0(freq$group,'_',freq$sex)
    freq$group_sex = factor(freq$group_sex, levels = c('control_F','trisomy8_F','control_M','trisomy8_M'))
  }
  freq$orig.ident = factor(freq$orig.ident, levels = c('HV-1', 'HV-2', 'HV-3','TRIAD-1','TRIAD-3','TRIAD-7', 'TRIAD-2', 'TRIAD-18','TRIAD-4'))
  return(freq)
}
# custom function to plot DE genes between Trisomy and Normal by clusters
plot_de_volcano_scatter = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Trisomy vs Normal') {
  # remove everything after "..." in gene names; bind_rows automatically added these to avoid duplicate rownames.
  df$genes = sub(pattern = '\\.\\.\\.*', replacement = '',x = rownames(df))
  # calculate log10 q value. cap log10 q value at 5 (which is 0.00001 in q val)
  df$p_val_adj_log10 = -log10(df$p_val_adj)
  df$p_val_adj_log10_capped = ifelse(df$p_val_adj_log10 < 5, df$p_val_adj_log10, 5)
  # generate summary statistics to show on caption
  minFDR <- -log10(maxq)
  sigs <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) >= minLFC
  sigLowFC <- df$p_val_adj_log10 >= minFDR & abs(df$avg_log2FC) < minLFC
  
  cnt_plus <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC >= minLFC)
  cnt_neg <- sum(df$p_val_adj_log10 >= minFDR & df$avg_log2FC <= -minLFC)
  if(showsub) {
    subtitle <- paste0(sum(sigs), " DE genes: ", cnt_plus, " Pos and ", cnt_neg, " Neg",
                       " (FDR<", maxq, ", log2FC>", minLFC, ")")
  } else {
    subtitle <- NULL
  }
  
  p1 = ggplot(df) +
    geom_point(aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5) + 
    geom_point(data=df[sigs,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigcol) +
    geom_point(data=df[sigLowFC,], aes(x=avg_log2FC, y=p_val_adj_log10_capped), size=0.5, col=sigLowFCcol) +
    ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("log10 FDR") +
    facet_wrap(.~Celltype_tested,nrow=round(length(unique(df$Celltype_tested))/2))+
    theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) + 
    geom_vline(xintercept=minLFC, col=sigcol, linetype="dashed", size=0.5) + 
    geom_vline(xintercept=-minLFC, col=sigcol, linetype="dashed", size=0.5) + 
    geom_hline(yintercept=minFDR, col=sigcol, linetype="dashed", size=0.5)
  
  df.s = subset(df, p_val_adj_log10 >minFDR & abs(avg_log2FC) > minLFC)
  df.s = df.s[order(df.s$avg_log2FC, decreasing = TRUE), ]
  df.s$genes = factor(df.s$genes,levels = unique(df.s$genes))
  p2 = ggplot(df.s,mapping = aes(x=genes,y=avg_log2FC,color=p_val_adj_log10_capped)) +
    geom_point() +
    facet_wrap(~ Celltype_tested, scales = 'free') +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    coord_flip() +
    ggtitle(title, subtitle=subtitle) + xlab("log2 Fold Change") + ylab("genes")
  return(list(p1,p2))
}

create_dt <- function(df,caption){
  datatable(df,
            caption = caption,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All")),
                           initComplete = JS(
    "function(settings, json) {",
    "$('caption').css({'font-size': '20px', 'font-family': 'Arial', 'font-weight': 'bold'});",
    "}")))
}

# function to set a column as rownames and remove that column from the df; default is setting the first column as the rowname
setrowname =  function(df,col = 1){
  if (is.numeric(col)) {
    # print('it is a number')
    rownames(df) = df[,col]
    df = df[,-col]
  } else {
    # print('it is a string')
    rownames(df) = df[,which(names(df) %in% c(col))]
    df = df[,-which(names(df) %in% c(col))]
  }
  return(df)
} 

# function to get colors from a palatte
getcol = function(group,pal = "Darjeeling1",shift_col_palatte=0){
  groupcol = as.numeric(factor(group))
  groupcol = wes_palette(pal)[groupcol+shift_col_palatte]
  return(groupcol)
}


# Volcano plot
plot_de_volcano_scatter3 = function(df, maxq = 0.1, minLFC=1, showsub = T, sigcol = 'red', sigLowFCcol = 'navyblue', title = 'Volcano plot',
                                    axislim,leftarrowlim,rightarrowlim) {
  # calculate log10 q value. 
  df$qval_log10 = -log10(df$`adj.P.Val`)
  # generate summary statistics to show on caption
  minFDR <- -log10(maxq)
  sigs <- df$qval_log10 >= minFDR & abs(df$logFC) >= minLFC
  sigLowFC <- df$qval_log10 >= minFDR & abs(df$logFC) < minLFC
  
  cnt_plus <- sum(df$qval_log10 >= minFDR & df$logFC >= minLFC)
  cnt_neg <- sum(df$qval_log10 >= minFDR & df$logFC <= -minLFC)
  if(showsub) {
    subtitle <- paste0(sum(sigs), " DEGs: ", cnt_plus, " + and ", cnt_neg, " -",
                       " (q<", maxq, ", log2FC>", minLFC, ")")
  } else {
    subtitle <- NULL
  }
  
  p1 = ggplot(df) +
    geom_point(aes(x=logFC, y=qval_log10, color = logFC, size = qval_log10)) + 
    geom_point(data =  df %>%
                 subset(qval_log10 >= minFDR) %>%
                 dplyr::arrange(desc(logFC)) %>%
                 dplyr::slice(1:5),
               aes(x = logFC, y = qval_log10,
                   # fill = log2FoldChange,
                   size = qval_log10),
               shape = 21, show.legend = F, color = "#000000") +
    geom_text_repel(data =  df %>%
                      subset(qval_log10 >= minFDR) %>%
                      dplyr::arrange(desc(logFC)) %>%
                      dplyr::slice(1:5),
                    aes(x = logFC, y = qval_log10,label = geneID),
                    box.padding = 0.5,
                    nudge_x = 0.5,
                    nudge_y = 0.2,
                    segment.curvature = -0.1,
                    segment.ncp = 3,
                    # segment.angle = 10,
                    direction = "y", 
                    hjust = "left"
    ) + 
    geom_point(data =  df %>%
                 subset(qval_log10 >= minFDR) %>%
                 dplyr::arrange(logFC) %>%
                 dplyr::slice(1:5),
               aes(x = logFC, y = qval_log10,
                   # fill = log2FoldChange,
                   size = qval_log10),
               shape = 21, show.legend = F, color = "#000000") +
    geom_text_repel(data =  df %>%
                      subset(qval_log10 >= minFDR) %>%
                      dplyr::arrange(logFC) %>%
                      dplyr::slice(1:5),
                    aes(x = logFC, y = qval_log10,label = geneID),
                    box.padding = 0.5,
                    nudge_x = 0.5,
                    nudge_y = 0.2,
                    segment.curvature = -0.1,
                    segment.ncp = 3,
                    # segment.angle = 10,
                    direction = "y", 
                    hjust = "right"
    ) + 
    scale_color_gradientn(colours = c("#3288bd", "#66c2a5","#ffffbf", "#f46d43", "#9e0142"),
                          values = seq(0, 1, 0.2)) +
    scale_fill_gradientn(colours = c("#3288bd", "#66c2a5","#ffffbf", "#f46d43", "#9e0142"),
                         values = seq(0, 1, 0.2)) +
    ggtitle(title, subtitle=subtitle) + xlab("log2 FC") + ylab("-log10 q value") +
    labs(color = 'log2 FC', size = '-log10 q') +
    theme(panel.grid = element_blank(),
          legend.background = element_roundrect(color = "#808080", linetype = 1),
          axis.text = element_text(size = 13, color = "#000000"),
          axis.title = element_text(size = 15),
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5)) + 
    geom_vline(xintercept=minLFC, linetype=2) + 
    geom_vline(xintercept=-minLFC, linetype=2) + 
    geom_hline(yintercept=minFDR, linetype=4) +
    coord_cartesian(clip = "off") + 
    annotation_custom(
      grob = grid::segmentsGrob(
        y0 = unit(-10, "pt"),
        y1 = unit(-10, "pt"),
        arrow = arrow(angle = 45, length = unit(.2, "cm"), ends = "first"),
        gp = grid::gpar(lwd = 3, col = "#74add1")
      ), 
      xmin = leftarrowlim[1], 
      xmax = leftarrowlim[2],
      ymin = leftarrowlim[3],
      ymax = leftarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::textGrob(
        label = "Down",
        gp = grid::gpar(col = "#74add1")
      ),
      xmin = leftarrowlim[1], 
      xmax = leftarrowlim[2],
      ymin = leftarrowlim[3],
      ymax = leftarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::segmentsGrob(
        y0 = unit(-10, "pt"),
        y1 = unit(-10, "pt"),
        arrow = arrow(angle = 45, length = unit(.2, "cm"), ends = "last"),
        gp = grid::gpar(lwd = 3, col = "#d73027")
      ), 
      xmin = rightarrowlim[1], 
      xmax = rightarrowlim[2],
      ymin = rightarrowlim[3],
      ymax = rightarrowlim[4]
    ) +
    annotation_custom(
      grob = grid::textGrob(
        label = "Up",
        gp = grid::gpar(col = "#d73027")
      ),
      xmin = rightarrowlim[1], 
      xmax = rightarrowlim[2],
      ymin = rightarrowlim[3],
      ymax = rightarrowlim[4]
    ) 
  if (length(axislim) == 4) {
    p1 = p1 + xlim(axislim[1], axislim[2]) + ylim(axislim[3], axislim[4])
  }
  return(p1)
}
```

## Load seurat object


```{r load pbmc, eval=T}
pbmc = readRDS('srt.obj.merged.clusterCuratedRefined.classifiedSet.myeloidInfoAdded.metadataIdentifierAdded.RDS')
pbmc.copy1 <- pbmc
#pbmc.copy1 -> pbmc
pbmc.metadata = pbmc@meta.data
# pbmc.mld2 = readRDS('pbmc.mld2.SexAgeAdded.RDS')
# pbmc.mld2.copy1 <- pbmc.mld2
#pbmc.mld2.copy1 -> pbmc.mld2
```

## The number of B and T cells in GEX data

```{r bt cell count}
btcellN = as.data.frame(table(pbmc@meta.data$orig.ident,pbmc@meta.data$curated_clusters4))
btcellN = subset(btcellN,Var2 %in% c('B memory','B Naive', 'Plasmablast B', 
                                     'CD4 T Memory',  'CD4 T Naive',
                                     'CD8 T Memory',  'CD8 T Naive',
                                      'gdT', 'MAIT'))

btcellN$CellType = ifelse(btcellN$Var2 %in% c('CD4 T Memory',  'CD4 T Naive',
                                     'CD8 T Memory',  'CD8 T Naive',
                                      'gdT', 'MAIT'), 'T cells',
                          ifelse(btcellN$Var2 %in% c('B memory','B Naive', 'Plasmablast B'), 'B cells','NA'))

btcellN.sum <- btcellN %>%
    group_by(Var1,CellType) %>%
    summarise(CellCount=sum(Freq))

create_dt(btcellN.sum,'The number of B and T cells in each sample')
```

## Load VDJ data

```{r load vdj,eval=FALSE}
# vdj b Immcantation reprocessed files
file_path = '/data/NCBR/projects/NCBR-323/Immcantation'

# read in the data
# specify the data types of non AIRR-C standard fields
# we assign integer type to the *_length fields
sampleList = c('TRIAD-18', 'TRIAD-7', 'TRIAD-2', 'TRIAD-3', 'TRIAD-4', 'TRIAD-1', 'HV-1', 'HV-3', 'HV-2')
bcr_data.merged = NULL
for (i in sampleList){
  bcr_data <- airr::read_rearrangement(file.path(file_path, paste0(i,"_BCR_data_sequences_igblast_db-pass.tsv_db-pass.tsv")),
                                     aux_types=c("v_germline_length"="i",
                                                 "d_germline_length"="i",
                                                 "j_germline_length"="i"))
  bcr_data$sample = i
  bcr_data.merged = bind_rows(bcr_data.merged,bcr_data)
}

bcr_data <- bcr_data.merged
bcr_data$cell_id_unique = paste0(bcr_data$sample,"_",bcr_data$cell_id)

saveRDS(bcr_data,'immdata.B.immcantation.rds')

```

```{r read vdj rds}
bcr_data = readRDS('immdata.B.immcantation.rds')
bcr_data.copy1 = bcr_data
```

## Metadata

```{r parse meta}
bcr_data$group = ifelse(bcr_data$sample %in% c('HV-1', 'HV-2', 'HV-3'), 'control', 'trisomy8')
```

# B cell VDJ analysis

## Preprocess BCR data to remove:

1. non-productive sequences

2. cells with multiple heavy chains

3. cells without heavy chains

```{r a1}
cat(paste("There are", nrow(bcr_data), "rows in the data before removing non-productive sequences.\n"))
# Remove non-productive sequences
bcr_data <- bcr_data %>% dplyr::filter(productive)
cat(paste("There are", nrow(bcr_data), "rows in the data after removing non-productive sequences.\n"))

# remove cells with multiple heavy chains
multi_heavy <- table(dplyr::filter(bcr_data, locus == "IGH")$cell_id_unique)
multi_heavy_cells <- names(multi_heavy)[multi_heavy > 1]
bcr_data <- dplyr::filter(bcr_data, !cell_id_unique %in% multi_heavy_cells)
cat(paste("There are", nrow(bcr_data),
          "rows in the data after filtering out cells with multiple heavy chains.\n"))

# Remove cells without heavy chains
# split cells by heavy and light chains
heavy_cells <- dplyr::filter(bcr_data, locus == "IGH")$cell_id_unique
light_cells <- dplyr::filter(bcr_data, locus == "IGK" | locus == "IGL")$cell_id_unique
no_heavy_cells <- light_cells[which(!light_cells %in% heavy_cells)]

bcr_data <- dplyr::filter(bcr_data, !cell_id_unique %in% no_heavy_cells)
cat(paste("There are", nrow(bcr_data), "rows in the data after filtering out
          cells without heavy chains."))
```

## Match cells between GEX and BCR data

```{r match}
#head(pbmc.metadata$Row.names)
#head(bcr_data$cell_id_unique)
# length(intersect(pbmc.metadata$Row.names,bcr_data$cell_id_unique))/length(unique(bcr_data$cell_id_unique))
bcr_GEX_VDJ_sharedCells = intersect(pbmc.metadata$Row.names,bcr_data$cell_id_unique)
#subset BCR to the cells appeared in both GEX and VDJ sets
bcr_data = subset(bcr_data,cell_id_unique %in% bcr_GEX_VDJ_sharedCells)
#merge cell annotation/grouping information from the GEX metadata
bcr_data = merge(bcr_data, select(pbmc.metadata,c(Row.names,trisomy,curated_clusters4,curated_clusters4_trisomy,curated_clusters4_group,curated_clusters4_group_trisomy)),
                 all.x = T,
                 by.x = 'cell_id_unique', by.y = 'Row.names')
cat(paste("There are", nrow(bcr_data), "rows in the data after matching with GEX data."))
cat(paste("There are", length(unique(bcr_data$cell_id_unique)), "VDJ-GEX shared cells across all samples."))
```


## Identify clonal threshold

We don't observe bimodal distribution from any of the samples, and therefore the prediction of clonal cutoff threshold failed.

Here I just use threshold = 0.1 for all samples. Most of the clonal lineages will only have one clonetype.

```{r clonal threshold}
# calculate cross subjects distribution of distance to nearest
dist_cross <- distToNearest(dplyr::filter(bcr_data, locus == "IGH"),
                                nproc = 1, cross = "sample")

ggplot(subset(dist_cross, !is.na(cross_dist_nearest)), 
             aes(x=cross_dist_nearest)) + 
    theme_bw() + 
    xlab("Cross-sample Hamming distance") + 
    ylab("Count") +
  labs(title = "Nearest-neighbor distance distribution",
       subtitle = "arbitrary cutoff = 0.1") +
    geom_histogram(color="white", binwidth=0.02) +
    geom_vline(xintercept=0.1, color="firebrick", linetype=2) +
    facet_grid(sample ~ ., scales="free_y")

```


```{r temp}
# calculate distribution of distance to nearest for each subject individually
sampleList = unique(bcr_data$sample)
for (i in sampleList){
  print(i)
  dist_ham <- distToNearest(dplyr::filter(dist_cross, locus == "IGH",
                                            sample == i), )
  output <- findThreshold(dist_ham$dist_nearest, method="density")
  threshold <- output@threshold
  
  g = ggplot(subset(dist_ham, !is.na(dist_nearest)),aes(x=dist_nearest,
   y = ..density..)) + 
     theme_bw() + 
     xlab("Hamming distance") + 
     ylab("Count") +
     scale_x_continuous(breaks=seq(0, 1, 0.1)) +
     geom_histogram(color="white", binwidth=0.02) +
     ggtitle(paste(i,"| predicted threshold:",threshold))+
     geom_histogram(
       aes(x=cross_dist_nearest,y = -..density..),
       color="white", binwidth=0.02,fill="black")+
     xlim(0,max(filter(dist_cross,
       !is.na(cross_dist_nearest))$cross_dist_nearest))+
     geom_vline(xintercept=0.1,color="grey")
   if(!is.na(threshold)){
       g = g + geom_vline(xintercept=threshold, color="firebrick", linetype=2)
   }
  print(g)

}
```

## Define clonal groups


hierarchicalClones clusters B receptor sequences based on junction region sequence similarity within partitions that share the same V gene, J gene, and junction length, thus allowing for ambiguous V or J gene annotations. By setting it up the cell_id parameter, hierarchicalClones will run in single-cell mode with paired-chain sequences. With only_heavy = TRUE and split_light = TRUE, grouping should be done by using IGH only and inferred clones should be split by the light/short chain (IGK and IGL) following heavy/long chain clustering.


```{r define clonal types}
clones = bcr_data %>%
    group_by(sample) %>%
    do(as.data.frame(
    hierarchicalClones(., 
                       threshold = 0.1, 
                       cell_id = "cell_id_unique", 
                       locus = "locus", #default
                       only_heavy = FALSE, 
                       split_light = TRUE, #default
                       cdr3 = TRUE, 
                       nproc = 16,
                       verbose = T, 
                       log = NULL, #default
                       summarize_clones = TRUE))) #default

clones$clone_id = paste0(clones$sample,"-",clones$clone_id)

```

## Create germlines

```{r create germlines}
file_path = '/data/NCBR/projects/NCBR-323/Immcantation'
references <- readIMGT(dir = paste0(file_path,"/human/vdj"))

clones_germline = createGermlines(clones,references,nproc = 16,verbose = T)
```

## Calculate V gene SHM frequency

```{r cal v shm}
# this is typically only done on heavy chains
results_heavy <- dplyr::filter(clones_germline, locus == "IGH")

# calculate SHM frequency in the V gene
data_mut <- shazam::observedMutations(results_heavy,
                              sequenceColumn = "sequence_alignment",
                              germlineColumn = "germline_alignment_d_mask",
                              regionDefinition = IMGT_V,
                              frequency = TRUE,
                              combine = TRUE,
                              nproc = 16)


```

### Distribution of median mutation frequency of clones


```{r dis med mut freq clones}
# calculate the median mutation frequency of a clone
mut_freq_clone <- data_mut %>%
                    dplyr::group_by(clone_id) %>%
                    dplyr::summarize(median_mut_freq = median(mu_freq))

ggplot(mut_freq_clone, aes(median_mut_freq)) +
  geom_histogram(binwidth = 0.005) +
  theme_bw() + theme(axis.title = element_text(size = 18))

```

### Distribution of mutation frequency of cells

#### All B cells

```{r dist mut freq by samples}
ggboxplot(data_mut, x = "sample", y = "mu_freq",
                     add = "jitter",  # Add jittered points
                     ylab = "Mutation frequency", xlab = "Sample",
                     caption = "All B cells in BCR VDJ data")
ggboxplot(data_mut, x = "group", y = "mu_freq",
                     add = "jitter",  # Add jittered points
                     ylab = "Mutation frequency", xlab = "Group",
                     caption = "All B cells in BCR VDJ data") +
  stat_compare_means(method = "wilcox.test")

```


#### Memory B cells

```{r dist mut freq by samples mem}
data_mut.Bmemory = subset(data_mut,curated_clusters4 == 'B memory')
caption = "B memory cells annotated by GEX data"
tmp = data_mut.Bmemory

ggboxplot(tmp, x = "sample", y = "mu_freq",
                     add = "jitter",  # Add jittered points
                     ylab = "Mutation frequency", xlab = "Sample",
                     caption = caption)
ggboxplot(tmp, x = "group", y = "mu_freq",
                     add = "jitter",  # Add jittered points
                     ylab = "Mutation frequency", xlab = "Group",
                     caption = caption) +
  stat_compare_means(method = "wilcox.test")


data_mut.Bmemory.summary <- data_mut.Bmemory %>%
  group_by(curated_clusters4, sample) %>%
  summarise(
    mean_mu_freq = mean(mu_freq),
    sd_mu_freq = sd(mu_freq)
  )

create_dt(data_mut.Bmemory.summary,'Mean and standard deviation of mutation frequency of B memory cells of each sample')

write.csv(data_mut.Bmemory.summary,'data_mut.Bmemory.summary.session29.csv')

```

